---
title: "Data Cleaning: Loneliness modulates self-reported empathy but not empathic multi-voxel neural response patterns after meditation training"
author: "Marla Dressel"
date: "02/02/2023"
output:
html_document:
df_print: paged
---
  
```{r, message = FALSE}
# import packages
library(readr)
library(tidyr) 
library(plyr)
library(dplyr)  
library(data.table) 
library(stringr)
library(purrr)
library(readxl)
```

# 1. Screener Survey (T1)

### 1.1. import data (Week 1, pre intervention)

```{r, message = FALSE}

# Define the path to the directory
path <- "/my/path/data" # CHANGE THIS PATH TO YOUR DIRECTORY!

# Construct the full file path and import qualtrics screener data (time point 1)
df_T1 <- read.csv(file.path(path, "raw/rawdata_T1_preIntervention.csv"), header = TRUE)


```

### 1.2. gross clean and initial exclusions

```{r, message = FALSE}

# Choose valid subject IDs
df_T1$id[df_T1$id == "002"] <- "ML002"  # id was put in incorrectly
df_T1 <- df_T1[grepl("ML", df_T1$id),]  # exclude empty id rows
df_T1 <- df_T1[grepl(1, df_T1$Finished),]  # exclude unfinished surveys
colnames(df_T1) <- gsub("id", "SubID", colnames(df_T1)) #rename subject number column 

# Move SubID to the first column
df_T1 <- df_T1 %>%
  dplyr::select(SubID, everything())

# remove non numerical prefix
df_T1$SubID <- gsub("ML", "", df_T1$SubID) 
df_T1$SubID <- as.numeric(df_T1$SubID) # remove ML and make numeric
df_T1 <- df_T1[order(df_T1$SubID),] # reorder according to SubID

# one subject was given the SubID of a previous subject, so we manually change it to their correct ID
df_T1$SubID[df_T1$EndDate == "2023-04-28 09:06:52"] <- 137
df_T1 <- df_T1[order(df_T1$SubID),]

# two people took the survey twice, so we discard their second attempt manually (SubID 140 and 184)
df_T1 <- df_T1[!df_T1$ResponseId == "R_31yFirmClvxCiuN",]
df_T1 <- df_T1[!df_T1$ResponseId == "R_3CO4gNO1MZ6ctJX",]
```

### 1.3. select variables of interest

```{r, message = FALSE}

# SCS-R
socialConnectedness <- df_T1 %>% dplyr::select(SubID, contains("SCS"), -matches("_T_")) # making new data frame with all columns that contain the string SCS
reverse_cols <- c("SCS_1", "SCS_2", "SCS_3", "SCS_4","SCS_5", "SCS_6", "SCS_8","SCS_10","SCS_15", "SCS_18") # which ones are reverse coded 
socialConnectedness <- data.frame(apply(socialConnectedness, 2, as.numeric))
socialConnectedness[, reverse_cols] <- 7 - socialConnectedness[, reverse_cols] # reverse code
df_long_socialConnectedness <- melt(setDT(socialConnectedness), id.vars = "SubID") # long format
df_long_socialConnectedness <- df_long_socialConnectedness %>%
  group_by(SubID)%>%
  summarise_at(vars(value), list(socialConnectedness_mean_T1 = mean)) # calculate mean per participant

df_ml_T1 <- df_long_socialConnectedness

# UCLA
loneliness <- df_T1 %>% dplyr::select(SubID, contains("UCLA"), -matches("_T_"))
reverse_cols <- c("UCLA_1", "UCLA_5", "UCLA_9", "UCLA_10","UCLA_15", "UCLA_16", "UCLA_19","UCLA_20")
loneliness <- data.frame(apply(loneliness, 2, as.numeric))
loneliness[, reverse_cols] <- 5 - loneliness[, reverse_cols] # reverse code
df_long_loneliness <- melt(setDT(loneliness), id.vars = "SubID") # long format
df_long_loneliness <- df_long_loneliness %>%
  group_by(SubID)%>%
  summarise_at(vars(value), list(loneliness_mean_T1 = mean)) # calculate mean per participant

df_ml_T1 <- left_join(df_ml_T1, df_long_loneliness, by = "SubID")

# IRI
traitEmpathyIRI <- df_T1 %>% dplyr::select(SubID, contains("IRI"), -matches("_T_"))
df_long_traitEmpathyIRI <- melt(setDT(traitEmpathyIRI), id.vars = "SubID") # long format
df_long_traitEmpathyIRI$value = as.numeric(df_long_traitEmpathyIRI$value) # numeric
df_long_traitEmpathyIRI <- df_long_traitEmpathyIRI %>%
  group_by(SubID)%>%
  summarise_at(vars(value), list(traitEmpathyIRI_mean_T1 = mean)) # calculate mean per participant

df_ml_T1 <- left_join(df_ml_T1, df_long_traitEmpathyIRI, by = "SubID")

#IRI subscales
#Fantasy
df_traitEmpathyIRI_fant <- traitEmpathyIRI %>% dplyr::select("SubID","IRI_1","IRI_5", "IRI_7", "IRI_12", "IRI_16", "IRI_23", "IRI_26")
df_long_traitEmpathyIRI_fant <- melt(setDT(df_traitEmpathyIRI_fant), id.vars = "SubID") # long format
df_long_traitEmpathyIRI_fant$value = as.numeric(df_long_traitEmpathyIRI_fant$value)
df_long_traitEmpathyIRI_fant <- df_long_traitEmpathyIRI_fant %>%
  group_by(SubID)%>%
  summarise_at(vars(value), list(traitEmpathyIRI_fant_mean_T1 = mean)) # calculate mean per participant

df_ml_T1 <- left_join(df_ml_T1, df_long_traitEmpathyIRI_fant, by = "SubID")

#Perspective-taking
df_traitEmpathyIRI_PT <- traitEmpathyIRI %>% dplyr::select("SubID","IRI_3","IRI_8", "IRI_11", "IRI_15", "IRI_21", "IRI_25", "IRI_28")
df_long_traitEmpathyIRI_PT <- melt(setDT(df_traitEmpathyIRI_PT), id.vars = "SubID") # long format
df_long_traitEmpathyIRI_PT$value = as.numeric(df_long_traitEmpathyIRI_PT$value)
df_long_traitEmpathyIRI_PT <- df_long_traitEmpathyIRI_PT %>%
  group_by(SubID)%>%
  summarise_at(vars(value), list(traitEmpathyIRI_PT_mean_T1 = mean)) # calculate mean per participant

df_ml_T1 <- left_join(df_ml_T1, df_long_traitEmpathyIRI_PT, by = "SubID")

#Empathic Concern
df_traitEmpathyIRI_EC <- traitEmpathyIRI %>% dplyr::select("SubID","IRI_2","IRI_4", "IRI_9", "IRI_14", "IRI_18", "IRI_20", "IRI_22")
df_long_traitEmpathyIRI_EC <- melt(setDT(df_traitEmpathyIRI_EC), id.vars = "SubID") # long format
df_long_traitEmpathyIRI_EC$value = as.numeric(df_long_traitEmpathyIRI_EC$value)
df_long_traitEmpathyIRI_EC <- df_long_traitEmpathyIRI_EC %>%
  group_by(SubID)%>%
  summarise_at(vars(value), list(traitEmpathyIRI_EC_mean_T1 = mean)) # calculate mean per participant

df_ml_T1 <- left_join(df_ml_T1, df_long_traitEmpathyIRI_EC, by = "SubID")

#Personal Distress
df_traitEmpathyIRI_PD <- traitEmpathyIRI %>% dplyr::select("SubID","IRI_6","IRI_10", "IRI_13", "IRI_17", "IRI_19", "IRI_24", "IRI_27")
df_long_traitEmpathyIRI_PD <- melt(setDT(df_traitEmpathyIRI_PD), id.vars = "SubID") # long format
df_long_traitEmpathyIRI_PD$value = as.numeric(df_long_traitEmpathyIRI_PD$value)
df_long_traitEmpathyIRI_PD <- df_long_traitEmpathyIRI_PD %>%
  group_by(SubID)%>%
  summarise_at(vars(value), list(traitEmpathyIRI_PD_mean_T1 = mean)) # calculate mean per participant

df_ml_T1 <- left_join(df_ml_T1, df_long_traitEmpathyIRI_PD, by = "SubID")

# IOS
inclusionOtherSelf <- df_T1 %>% dplyr::select(SubID, contains("IOS"))
inclusionOtherSelf$IOS_1b <- as.numeric(inclusionOtherSelf$IOS_1b)
inclusionOtherSelf$IOS_2b <- as.numeric(inclusionOtherSelf$IOS_2b)
inclusionOtherSelf$IOS_3b <- as.numeric(inclusionOtherSelf$IOS_3b)
inclusionOtherSelf$IOS_4b <- as.numeric(inclusionOtherSelf$IOS_4b)
inclusionOtherSelf$IOS_5 <- as.numeric(inclusionOtherSelf$IOS_5)
inclusionOtherSelf$inclusionOtherSelf_mean <- rowMeans(inclusionOtherSelf[,c("IOS_1b", "IOS_2b", "IOS_3b", "IOS_4b", "IOS_5")])
inclusionOtherSelf$inclusionOtherSelf_mean_nostranger <- rowMeans(inclusionOtherSelf[,c("IOS_1b", "IOS_2b", "IOS_3b", "IOS_4b")]) # IOS but not for stranger 
colnames(inclusionOtherSelf) <- c("SubID", "IOS_immediateFamily_T1", "IOS_immediateFamilyNum_T1", "IOS_extendedFamily_T1", "IOS_extendedFamilyNum_T1", "IOS_friend_T1", "IOS_friendNum_T1", "IOS_acquaintance_T1", "IOS_acquaintanceNum_T1", "IOS_strangerNum_T1", "inclusionOtherSelf_mean_T1", "inclusionOtherSelf_mean_nostranger_T1")

df_ml_T1 <- left_join(df_ml_T1, inclusionOtherSelf, by = "SubID")

# Cantrill wellbeing Ladder
wellbeing <- df_T1 %>% dplyr::select(SubID, contains("LADDER"))
wellbeing$LADDER_PRESENT <- as.numeric(as.character(wellbeing$LADDER_PRESENT))
wellbeing$LADDER_FUTURE  <- as.numeric(as.character(wellbeing$LADDER_FUTURE))
wellbeing$wellbeing_mean_T1 <- rowMeans(wellbeing[, c("LADDER_PRESENT", "LADDER_FUTURE")], na.rm = TRUE)
colnames(wellbeing) <- c("SubID", "wellbeingPresent_T1", "wellbeingFuture_T1", "wellbeing_mean_T1")

df_ml_T1 <- left_join(df_ml_T1, wellbeing, by = "SubID")

```

### 1.4. prior meditation experience

```{r, message = FALSE}
## prior meditation experience
priorMedExp <- df_T1 %>% dplyr::select(SubID, matches("MEDITATION")) 
colnames(priorMedExp) <- c("SubID","priorMedExpClosed", "priorMedExpOpen", "priorMedExpHours")

df_ml_T1 <- left_join(df_ml_T1, priorMedExp, by = "SubID")

```

### 1.5. demographics 

```{r, message = FALSE}

# age
df_T1$AGE[df_T1$SubID == 122] <- 64 # one participant (122) put birth year and month (march 1959) but not age 

age <- df_T1 %>% dplyr::select(SubID, contains("AGE"), -matches("LANGUAGE"), -matches("PAGE"))
age <- age %>%
  mutate(agefull = case_when(
    AGE <= 39 ~ "young adult",
    AGE >= 40 & AGE <= 59 ~ "middle adult",
    AGE >= 60 ~ "old adult"
  ))
colnames(age) <- c("SubID", "age", "agefull")

df_ml_T1 <- left_join(df_ml_T1, age, by = "SubID")

# single column demographics
demo <- df_T1 %>% dplyr::select(SubID, contains("BIRTH"), contains("PRONOUNS"), contains("USLIVED"), 
                                     contains("LANGUAGE"),  contains("YRSENGLISH"), 
                                     contains("ENGPROF"), contains("JOBTITLE"), contains("PUBLIC ASSISTANCE"), 
                                     contains("HOUSENUM"), contains("householdCOMP"), contains("POL"), 
                                     contains("EVANG"), contains("SPIRT"))

colnames(demo) <- c("SubID", "birth", "pronouns", 
                   "yearslivedUS", "UserLang", "firstLangEng", 
                   "yearsspeakingEng", "EnglishProficiency", 
                   "jobtitle", "householdNum", "householdComp1", 
                   "householdComp2", "householdComp3", "householdComp4", 
                   "politicalIssues", "socialIssues", "economicIssues", 
                   "evangelic", "spirituality")

df_ml_T1 <- left_join(df_ml_T1, demo, by = "SubID")

# gender
gender <- df_T1 %>% dplyr::select(SubID, contains("GENDER"), -matches("GENDER_4_TEXT"))
gender$GENDER[gender$GENDER == 1] <- "female"
gender$GENDER[gender$GENDER == 2] <- "male"
gender$GENDER[gender$GENDER == 3 | gender$GENDER == 4] <- "other"
colnames(gender) <- c("SubID", "gender")

df_ml_T1 <- left_join(df_ml_T1, gender, by = "SubID")

# ethnicity
ethnicity <- df_T1 %>% dplyr::select(SubID, contains("RACEETHN"))
ethnicity$RACEETHN[ethnicity$RACEETHN == 1] <- "Hispanic"
ethnicity$RACEETHN[ethnicity$RACEETHN == 2] <- "White"
ethnicity$RACEETHN[ethnicity$RACEETHN == 3] <- "Black"
ethnicity$RACEETHN[ethnicity$RACEETHN == 4] <- "Native American"
ethnicity$RACEETHN[ethnicity$RACEETHN == 5] <- "Asian"
ethnicity$RACEETHN[ethnicity$RACEETHN == 6] <- "Pacific Islander"
ethnicity$RACEETHN[ethnicity$RACEETHN == 7] <- "Other"
colnames(ethnicity) <- c("SubID", "ethnicity", "ethnicityDescribe")

df_ml_T1 <- left_join(df_ml_T1, ethnicity, by = "SubID")

# education
educationreduced <- df_T1 %>% # Initial mapping with grosser categories
  dplyr::select(SubID, contains("EDUCATION")) %>%
  mutate(
    EDUCATION = case_when(
      EDUCATION == 1 | EDUCATION == 2 | EDUCATION == 3 ~ "NoCollegeDegree",
      EDUCATION %in% 4:7 ~ "YesCollegeDegree",
      TRUE ~ EDUCATION
    )
  )
colnames(educationreduced) <- c("SubID", "education")

df_ml_T1 <- left_join(df_ml_T1, educationreduced, by = "SubID")

educationfull <- df_T1 %>% # More detailed mapping
  dplyr::select(SubID, contains("EDUCATION")) %>%
  mutate(
    EDUCATIONFULL = case_when(
      EDUCATION == 1 ~ "Less than high school",
      EDUCATION == 2 ~ "High school graduate or G.E.D.",
      EDUCATION == 3 ~ "Some college",
      EDUCATION == 4 ~ "Two-year college degree (A.A., A.S.)",
      EDUCATION == 5 ~ "Four-year college degree (B.A., B.S.)",
      EDUCATION == 6 ~ "Masters degree (for example, M.B.A., M.A., M.S.)",
      EDUCATION == 7 ~ "Doctoral or professional degree (for example, Ph.D., M.D., J.D., D.V.M., D.D.M.)",
      TRUE ~ EDUCATION  # This preserves the original value if none of the above groups are met
    )
  )

colnames(educationfull) <- c("SubID", "education", "educationfull")
educationfull <- educationfull[-2] 

df_ml_T1 <- left_join(df_ml_T1, educationfull, by = "SubID")

# employment
employmentreduced <- df_T1 %>%
  dplyr::select(SubID, contains("OCCSTAT")) %>%
  mutate(
    OCCSTAT = case_when(
      OCCSTAT == 1 | OCCSTAT == 2 ~ "Employed",
      OCCSTAT %in% c(3, 4, 7, 13) ~ "Other",
      OCCSTAT %in% c(5, 6) ~ "Student",
      OCCSTAT %in% c(8, 9, 10, 11, 12) ~ "Unemployed",
      TRUE ~ OCCSTAT
    )
  )
colnames(employmentreduced) <- c("SubID", "employment", "employmentDescribe")

df_ml_T1 <- left_join(df_ml_T1, employmentreduced, by = "SubID")

employmentfull <- df_T1 %>%
  dplyr::select(SubID, contains("OCCSTAT"), -matches("TEXT")) %>%
  mutate(
    OCCSTATFULL = case_when(
      OCCSTAT == 1 ~ "Employed full-time for pay",
      OCCSTAT == 2 ~ "Employed part-time for pay",
      OCCSTAT == 3 ~ "Homemaker",
      OCCSTAT == 4 ~ "Retired",
      OCCSTAT == 5 ~ "Full-time student",
      OCCSTAT == 6 ~ "Part-time student",
      OCCSTAT == 7 ~ "Leave of absence for medical reasons",
      OCCSTAT == 8 ~ "Unemployed less than 6 months, do not expect to work",
      OCCSTAT == 9 ~ "Unemployed more than 6 months, do not expect to work",
      OCCSTAT == 10 ~ "Unemployed less than 6 months, expects to work",
      OCCSTAT == 11 ~ "Unemployed more than 6 months, expects to work",
      OCCSTAT == 12 ~ "Laid off",
      OCCSTAT == 13 ~ "Other",
      TRUE ~ OCCSTAT
    )
  )
colnames(employmentfull) <- c("SubID", "employment", "employmentfull")
employmentfull <- employmentfull[-2] 

df_ml_T1 <- left_join(df_ml_T1, employmentfull, by = "SubID")

# income
incomereduced <- df_T1 %>%
  dplyr::select(SubID, contains("INCOME")) %>%
  mutate(
    INCOME = case_when(
      INCOME %in% 1:6 ~ "Lower",
      INCOME %in% 7:9 ~ "Upper",
      INCOME == 10 ~ "Dont_know",
      TRUE ~ INCOME
    )
  )
colnames(incomereduced) <- c("SubID", "income")

df_ml_T1 <- left_join(df_ml_T1, incomereduced, by = "SubID")


incomefull <- df_T1 %>%
  dplyr::select(SubID, contains("INCOME")) %>%
  mutate(
    INCOMEFULL = case_when(
      INCOME == 1 ~ "Under $5,000",
      INCOME == 2 ~ "$5,000 - $9,999",
      INCOME == 3 ~ "$10,000 - $14,999",
      INCOME == 4 ~ "$15,000 - $24,999",
      INCOME == 5 ~ "$25,000 - $39,999",
      INCOME == 6 ~ "$40,000 - $59,999",
      INCOME == 7 ~ "$60,000 - $89,999",
      INCOME == 8 ~ "$90,000 - $179,999",
      INCOME == 9 ~ "Over $180,000",
      INCOME == 10 ~ "Dont know",
      TRUE ~ INCOME
    )
  )
colnames(incomefull) <- c("SubID", "income", "incomefull")
incomefull <- incomefull[-2]

df_ml_T1 <- left_join(df_ml_T1, incomefull, by = "SubID")

# religion
religion <- df_T1 %>% dplyr::select(SubID, contains("REL"), -matches("RELAFF_11_TEXT"))
religion$RELAFF[religion$RELAFF == 1] <- "Protestant"
religion$RELAFF[religion$RELAFF == 2] <- "Roman Catholic"
religion$RELAFF[religion$RELAFF == 3] <- "Mormon"
religion$RELAFF[religion$RELAFF == 4] <- "Orthodox"
religion$RELAFF[religion$RELAFF == 5] <- "Jewish"
religion$RELAFF[religion$RELAFF == 6] <- "Muslim"
religion$RELAFF[religion$RELAFF == 7] <- "Buddhist"
religion$RELAFF[religion$RELAFF == 8] <- "Hindu"
religion$RELAFF[religion$RELAFF == 9] <- "Atheist"
religion$RELAFF[religion$RELAFF == 10] <- "Agnostic"
religion$RELAFF[religion$RELAFF == 11] <- "Other"
religion$RELAFF[religion$RELAFF == 12] <- "Nothing"

colnames(religion) <- c("SubID", "religionAffiliation", "religionStrength")

df_ml_T1 <- left_join(df_ml_T1, religion, by = "SubID")

handedness <- df_T1 %>% dplyr::select(SubID, contains("HAND"))
handedness$HAND[handedness$HAND == 1] <- "Right"
handedness$HAND[handedness$HAND == 2] <- "Left"
handedness$HAND[handedness$HAND == 3] <- "Both"

colnames(handedness) <- c("SubID", "handedness")

df_ml_T1 <- left_join(df_ml_T1, handedness, by = "SubID")

# Generate dummy variables "Female", "Male", and "Other"
df_ml_T1$Female <- 0
df_ml_T1$Female[df_ml_T1$gender == "female"] <- 1

df_ml_T1$Male <- 0
df_ml_T1$Male[df_ml_T1$gender == "male"] <- 1

df_ml_T1$Other_gender <- 0
df_ml_T1$Other_gender[df_ml_T1$gender == "other"] <- 1

#education dummys
df_ml_T1$NoCollegeDegree <- 0
df_ml_T1$NoCollegeDegree[grepl("NoCollegeDegree", df_ml_T1$education)] <- 1

df_ml_T1$YesCollegeDegree  <- 0
df_ml_T1$YesCollegeDegree [grepl("YesCollegeDegree", df_ml_T1$education)] <- 1

#income dummy 
df_ml_T1$Lower <- 0
df_ml_T1$Lower[grepl("Lower", df_ml_T1$income)] <- 1

df_ml_T1$Upper <- 0
df_ml_T1$Upper[grepl("Upper", df_ml_T1$income)] <- 1

df_ml_T1$Dont_know <- 0
df_ml_T1$Dont_know[grepl("Dont_know", df_ml_T1$income)] <- 1

# employment dummy
df_ml_T1$Employed <- 0
df_ml_T1$Employed[grepl("Employed", df_ml_T1$employment)] <- 1

df_ml_T1$Unemployed <- 0
df_ml_T1$Unemployed[grepl("Unemployed", df_ml_T1$employment)] <- 1

df_ml_T1$Student <- 0
df_ml_T1$Student[grepl("Student", df_ml_T1$employment)] <- 1

df_ml_T1$Other_employment <- 0
df_ml_T1$Other_employment[grepl("Other", df_ml_T1$employment)] <- 1

# Generate dummy variables for ethnicity
df_ml_T1$White <- as.integer(df_ml_T1$ethnicity == "White") #count as white even though also said asian
df_ml_T1$Black <- as.integer(df_ml_T1$ethnicity == "Black") #count as black even though also said other
df_ml_T1$Hispanic <- as.integer(df_ml_T1$ethnicity == "Hispanic") #count as hispanic even though also said white
df_ml_T1$OtherRace <- as.integer(df_ml_T1$ethnicity == "Other" | df_ml_T1$ethnicity == "Pacific Islander" | df_ml_T1$ethnicity == "Native American" | df_ml_T1$ethnicity == "2,3,4" | df_ml_T1$ethnicity == "2,7" | df_ml_T1$ethnicity == "2,5" | df_ml_T1$ethnicity == "3,7" | df_ml_T1$ethnicity == "3,4" | df_ml_T1$ethnicity == "2,3" | df_ml_T1$ethnicity == "1,2" | df_ml_T1$ethnicity == "1,2,4" | df_ml_T1$ethnicity == "1,7") # count as other 
df_ml_T1$Asian <- as.integer(df_ml_T1$ethnicity == "Asian") #count as Asian even though also said White

# make sure to only have one ethnicity by the end for simplicity
df_ml_T1 <- df_ml_T1 %>%
  mutate(
    ethnicity = case_when(
      str_detect(ethnicity, "1") ~ "Hispanic",
      str_detect(ethnicity, "5") ~ "Asian",
      str_detect(ethnicity, "3") ~ "Black",
      str_detect(ethnicity, "Native") ~ "Other",
      str_detect(ethnicity, "Pacific") ~ "Other",
      TRUE ~ ethnicity
    )
  )

# make sure to only one occupational status
df_ml_T1 <- df_ml_T1 %>%
  mutate(
    employment = case_when(
      str_detect(employment, "1|2") ~ "Employed",
      str_detect(employment, "5|6") ~ "Student",
      str_detect(employment, "3|4|7") ~ "Other",
      str_detect(employment, "8|9|10|11|12") ~ "Unemployed",
      TRUE ~ employment
    )
  )

# make sure you make dummies for gender as well
df_ml_T1$Female <- ifelse(df_ml_T1$gender=="female", 1, 0)
df_ml_T1$Male <- ifelse(df_ml_T1$gender=="male", 1, 0)
df_ml_T1$Other <- ifelse(df_ml_T1$gender=="other", 1, 0)


```

### 1.6. exclusion variables and saving the file as csv

```{r, message = FALSE}
# control questions
control <- df_T1 %>% dplyr::select(SubID, contains("RECRUIT"), contains("OPEN"), contains("COMMIT"),
                                        contains("ATTN_1"), # in SCS (strongly agree - 6)
                                        contains("ATTN_2"), # in IRI (does not describe me well - A)
                                        contains("ATTN_3")) # in UCLA (rarely - 2)

colnames(control) <- c("SubID", "recruited", "openQuestion", "commitmentRequest", "attentionCheck1_T1", "attentionCheck2_T1", "attentionCheck3_T1")

df_ml_T1 <- left_join(df_ml_T1, control, by = "SubID")

write.csv(df_ml_T1, file.path(path, "clean/cleandata_T1.csv"))


```

# 2. 4 Week Survey (T2)

### 2.1. import data from T2 (Week 4, post intervention)

```{r, message = FALSE}

# import data from four week survey
df_T2 <- read.csv(file.path(path, "raw/rawdata_T2_4weeks_postIntervention.csv"), header = T) 

# SubID 009 completed the survey twice, so we only count the first time they did it 
df_T2 <- df_T2[!df_T2$ResponseId == "R_28GJqKvSCZZ79NN",]

```

### 2.2. gross clean and initial exclusions

```{r, message = FALSE}

# gross clean
df_T2 <- df_T2[grepl("ML", df_T2$id),] # exclude empty id rows
df_T2 <- df_T2[grepl(1, df_T2$Finished),] # exclude unfinished surveys
colnames(df_T2) <- gsub("id", "SubID", colnames(df_T2))
df_T2 <- df_T2 %>%
  dplyr::select(SubID, everything()) # move SubID to first row
df_T2$SubID <- gsub("ML", "", df_T2$SubID) 
df_T2$SubID <- as.numeric(df_T2$SubID) # remove prefix ML and make numeric
df_T2 <- df_T2[order(df_T2$SubID),] # reorder according to SubID

# SubID 196 filled out week4 twice, so we remove the second attempt 
df_T2 <- df_T2[!df_T2$ResponseId == "R_xcnqvWuCDanFimZ",]

```

### 2.3. select variables of interest

```{r, message = FALSE}

# SCS
socialConnectedness <- df_T2 %>% dplyr::select(SubID, contains("SCS2"), -matches("_T_"))
reverse_cols <- c("SCS2_1", "SCS2_2", "SCS2_3", "SCS2_4","SCS2_5", "SCS2_6", "SCS2_8","SCS2_10","SCS2_15", "SCS2_18")
socialConnectedness <- data.frame(apply(socialConnectedness, 2, as.numeric))
socialConnectedness[,reverse_cols] <- 7 - socialConnectedness[,reverse_cols] # reverse code
df_long_socialConnectedness <- melt(setDT(socialConnectedness), id.vars = "SubID") # long format
df_long_socialConnectedness <- df_long_socialConnectedness %>%
  group_by(SubID)%>%
  summarise_at(vars(value), list(socialConnectedness_mean_T2 = mean)) # calculate mean per participant

df_ml_T2 <- df_long_socialConnectedness

# UCLA
loneliness <- df_T2 %>% dplyr::select(SubID, contains("UCLA2"), -matches("_T_"))
reverse_cols <- c("UCLA2_1", "UCLA2_5", "UCLA2_9", "UCLA2_10","UCLA2_15", "UCLA2_16", "UCLA2_19","UCLA2_20")
loneliness <- data.frame(apply(loneliness, 2, as.numeric))
loneliness[, reverse_cols] <- 5 - loneliness[, reverse_cols] # reverse code
df_long_loneliness <- melt(setDT(loneliness), id.vars = "SubID") # long format
df_long_loneliness <- df_long_loneliness %>%
  group_by(SubID)%>%
  summarise_at(vars(value), list(loneliness_mean_T2 = mean)) # calculate mean per participant

df_ml_T2 <- left_join(df_ml_T2, df_long_loneliness, by = "SubID")


# IRI
traitEmpathyIRI <- df_T2 %>% dplyr::select(SubID, contains("IRI2"), -matches("_T_"))
df_long_traitEmpathyIRI <- melt(setDT(traitEmpathyIRI), id.vars = "SubID") # long format
df_long_traitEmpathyIRI$value = as.numeric(df_long_traitEmpathyIRI$value) # numeric
df_long_traitEmpathyIRI <- df_long_traitEmpathyIRI %>%
  group_by(SubID)%>%
  summarise_at(vars(value), list(traitEmpathyIRI_mean_T2 = mean)) # calculate mean per participant

df_ml_T2 <- left_join(df_ml_T2, df_long_traitEmpathyIRI, by = "SubID")

#IRI subscales
#Fantasy
df_traitEmpathyIRI_fant <- traitEmpathyIRI %>% dplyr::select("SubID","IRI2_1","IRI2_5", "IRI2_7", "IRI2_12", "IRI2_16", "IRI2_23", "IRI2_26")
df_long_traitEmpathyIRI_fant <- melt(setDT(df_traitEmpathyIRI_fant), id.vars = "SubID") # long format
df_long_traitEmpathyIRI_fant$value = as.numeric(df_long_traitEmpathyIRI_fant$value)
df_long_traitEmpathyIRI_fant <- df_long_traitEmpathyIRI_fant %>%
  group_by(SubID)%>%
  summarise_at(vars(value), list(traitEmpathyIRI_fant_mean_T2 = mean)) # calculate mean per participant

df_ml_T2 <- left_join(df_ml_T2, df_long_traitEmpathyIRI_fant, by = "SubID")

#Perspective-taking
df_traitEmpathyIRI_PT <- traitEmpathyIRI %>% dplyr::select("SubID","IRI2_3","IRI2_8", "IRI2_11", "IRI2_15", "IRI2_21", "IRI2_25", "IRI2_28")
df_long_traitEmpathyIRI_PT <- melt(setDT(df_traitEmpathyIRI_PT), id.vars = "SubID") # long format
df_long_traitEmpathyIRI_PT$value = as.numeric(df_long_traitEmpathyIRI_PT$value)
df_long_traitEmpathyIRI_PT <- df_long_traitEmpathyIRI_PT %>%
  group_by(SubID)%>%
  summarise_at(vars(value), list(traitEmpathyIRI_PT_mean_T2 = mean)) # calculate mean per participant

df_ml_T2 <- left_join(df_ml_T2, df_long_traitEmpathyIRI_PT, by = "SubID")

#Empathic Concern
df_traitEmpathyIRI_EC <- traitEmpathyIRI %>% dplyr::select("SubID","IRI2_2","IRI2_4", "IRI2_9", "IRI2_14", "IRI2_18", "IRI2_20", "IRI2_22")
df_long_traitEmpathyIRI_EC <- melt(setDT(df_traitEmpathyIRI_EC), id.vars = "SubID") # long format
df_long_traitEmpathyIRI_EC$value = as.numeric(df_long_traitEmpathyIRI_EC$value)
df_long_traitEmpathyIRI_EC <- df_long_traitEmpathyIRI_EC %>%
  group_by(SubID)%>%
  summarise_at(vars(value), list(traitEmpathyIRI_EC_mean_T2 = mean)) # calculate mean per participant

df_ml_T2 <- left_join(df_ml_T2, df_long_traitEmpathyIRI_EC, by = "SubID")

#Personal Distress
df_traitEmpathyIRI_PD <- traitEmpathyIRI %>% dplyr::select("SubID","IRI2_6","IRI2_10", "IRI2_13", "IRI2_17", "IRI2_19", "IRI2_24", "IRI2_27")
df_long_traitEmpathyIRI_PD <- melt(setDT(df_traitEmpathyIRI_PD), id.vars = "SubID") # long format
df_long_traitEmpathyIRI_PD$value = as.numeric(df_long_traitEmpathyIRI_PD$value)
df_long_traitEmpathyIRI_PD <- df_long_traitEmpathyIRI_PD %>%
  group_by(SubID)%>%
  summarise_at(vars(value), list(traitEmpathyIRI_PD_mean_T2 = mean)) # calculate mean per participant

df_ml_T2 <- left_join(df_ml_T2, df_long_traitEmpathyIRI_PD, by = "SubID")

# IOS
inclusionOtherSelf <- df_T2 %>% dplyr::select(SubID, contains("IOS2"))
inclusionOtherSelf$IOS2_1b <- as.numeric(inclusionOtherSelf$IOS2_1b)
inclusionOtherSelf$IOS2_2b <- as.numeric(inclusionOtherSelf$IOS2_2b)
inclusionOtherSelf$IOS2_3b <- as.numeric(inclusionOtherSelf$IOS2_3b)
inclusionOtherSelf$IOS2_4b <- as.numeric(inclusionOtherSelf$IOS2_4b)
inclusionOtherSelf$IOS2_5 <- as.numeric(inclusionOtherSelf$IOS2_5)
inclusionOtherSelf$inclusionOtherSelf_mean <- rowMeans(inclusionOtherSelf[,c("IOS2_1b", "IOS2_2b", "IOS2_3b", "IOS2_4b", "IOS2_5")])
inclusionOtherSelf$inclusionOtherSelf_mean_nostranger <- rowMeans(inclusionOtherSelf[,c("IOS2_1b", "IOS2_2b", "IOS2_3b", "IOS2_4b")])
colnames(inclusionOtherSelf) <- c("SubID", "IOS_immediateFamily_T2", "IOS_immediateFamilyNum_T2", "IOS_extendedFamily_T2", "IOS_extendedFamilyNum_T2", "IOS_friend_T2", "IOS_friendNum_T2", "IOS_acquaintance_T2", "IOS_acquaintanceNum_T2", "IOS_strangerNum_T2", "inclusionOtherSelf_mean_T2", "inclusionOtherSelf_mean_nostranger_T2")

df_ml_T2 <- left_join(df_ml_T2, inclusionOtherSelf, by = "SubID")

# Cantrill wellbeing ladder
wellbeing <- df_T2 %>% dplyr::select(SubID, contains("LADDER"))
wellbeing$LADDER_PRESENT2 <- as.numeric(as.character(wellbeing$LADDER_PRESENT2))
wellbeing$LADDER_FUTURE2  <- as.numeric(as.character(wellbeing$LADDER_FUTURE2))
wellbeing$wellbeing_mean_T2 <- rowMeans(wellbeing[, c("LADDER_PRESENT2", "LADDER_FUTURE2")], na.rm = TRUE)
colnames(wellbeing) <- c("SubID", "wellbeingPresent_T2", "wellbeingFuture_T2", "wellbeing_mean_T2")

df_ml_T2 <- left_join(df_ml_T2, wellbeing, by = "SubID")

```


### 2.4. exclusion variables and saving the file as csv

```{r, message = FALSE}

# control questions
attention <- df_T2 %>% dplyr::select(SubID, contains("ATTN2_1"), # in SCS (strongly agree - 6)
                                                contains("ATTN2_2"), # in IRI (does not describe me well - A)
                                                contains("ATTN2_3")) # in UCLA (rarely - 2)
colnames(attention) <- c("SubID", "attentionCheck1_T2", "attentionCheck2_T2", "attentionCheck3_T2")

df_ml_T2 <- left_join(df_ml_T2, attention, by = "SubID")

# save the file to csv 
write.csv(df_ml_T1, file.path(path, "clean/cleandata_T2.csv"))


```

# 3. Scan Day

### 3.1. Scan Day Survey 

#### 3.1.1. import data from the scanday qualtrics survey (the survey they completed after the scan)

```{r, message = FALSE}
# import data from the scan day survey 
df_scanSurvey <- read.csv(file.path(path, "raw/rawdata_scandaySurvey.csv"), header = T) 

```

#### 3.1.2. gross clean and initial exclusions

```{r, message = FALSE}
# gross clean
df_scanSurvey <- df_scanSurvey[grepl("ML", df_scanSurvey$id),] # exclude empty id rows
df_scanSurvey <- df_scanSurvey[grepl(1, df_scanSurvey$Finished),] # exclude unfinished surveys
colnames(df_scanSurvey) <- gsub("id", "SubID", colnames(df_scanSurvey))
df_scanSurvey <- df_scanSurvey %>%
  dplyr::select(SubID, everything()) # move SubID to first row
df_scanSurvey$SubID <- gsub("ML", "", df_scanSurvey$SubID) 
df_scanSurvey$SubID <- as.numeric(df_scanSurvey$SubID) # remove ML prefix and make numeric
df_scanSurvey <- df_scanSurvey[order(df_scanSurvey$SubID),] # reorder according to SubID

# SubID 9 completed the study twice, so we remove the second attempt (this person gets excluded anyways for falling asleep during the scan)
df_scanSurvey <- df_scanSurvey[!df_scanSurvey$ResponseId == "R_24dbzIIjZohhm4j",]

```

#### 3.1.3. select variables of interest 

```{r, message = FALSE}

# subjective ratings, based on which we make state empathy variable later on 
subjectiveRatings <- df_scanSurvey %>% dplyr::select(SubID, contains("PE"), -matches("_T_"))
colnames(subjectiveRatings) <- c("SubID", "self_fearOfPain", "self_fearNoPain", "self_unpleasantPain", "self_unpleasantNoPain", "other_fearOfPain", "other_fearNoPain", "other_unpleasantPain", "other_unpleasantNoPain")

# calculate state empathy by taking the difference between observed and experienced rating and then taking the absolute (so direction does not matter) value and then taking the inverse because higher difference between you + me = less similarity aka empathy. multiplying stateEmp_fear (the difference between observed and experienced self-reported empathy) by -1 effectively reverses its interpretation. Originally, higher values indicated greater differences between observing and experiencing pain. By inverting the scale (multiplying by -1), higher values now indicate greater similarity between the two conditions.
subjectiveRatings$stateEmp_fear <- abs(abs(as.numeric(subjectiveRatings$self_fearOfPain) - as.numeric(subjectiveRatings$self_fearNoPain)) - abs(as.numeric(subjectiveRatings$other_fearOfPain) - as.numeric(subjectiveRatings$self_fearNoPain)))

subjectiveRatings$stateEmp_painUnpleasant <- abs(abs(as.numeric(subjectiveRatings$self_unpleasantPain) - as.numeric(subjectiveRatings$self_unpleasantNoPain)) - abs(as.numeric(subjectiveRatings$other_unpleasantPain) - as.numeric(subjectiveRatings$self_unpleasantNoPain)))

subjectiveRatings$stateEmp_fear <- subjectiveRatings$stateEmp_fear * (-1)
subjectiveRatings$stateEmp_painUnpleasant <- subjectiveRatings$stateEmp_painUnpleasant * (-1)

df_ml_scanSurvey <- left_join(df_scanSurvey, subjectiveRatings, by = "SubID")

# IOS
inclusionOtherSelf <- df_scanSurvey %>% dplyr::select(SubID, contains("IOS"))
colnames(inclusionOtherSelf) = c("SubID", "IOS_studyPartner")

df_ml_scanSurvey <- left_join(df_ml_scanSurvey, inclusionOtherSelf, by = "SubID")

# study partner impressions 
studyPartner <- df_scanSurvey %>% dplyr::select(SubID, contains("Q590"), -matches("_T_"))
colnames(studyPartner) <- c("SubID", "studyPartnerLiking", "studyPartnerReal")

df_ml_scanSurvey <- left_join(df_ml_scanSurvey, studyPartner, by = "SubID")


```

#### 3.1.4. exclusion variables and saving the file as csv

```{r, message = FALSE}

# control variables 
df_control <- df_ml_scanSurvey %>% dplyr::select(SubID, Q595, Q589, Q591, ATTNS_2)
colnames(df_control) <- c("SubID", "medication", "guessStudy", "anythingElse", "attentionCheck_scanday")

df_ml_scanSurvey <- left_join(df_ml_scanSurvey, df_control, by = "SubID")

# write to csv
write.csv(df_ml_scanSurvey, file.path(path, "clean/cleandata_scandaySurvey.csv"))

```

### 3.2. Scan Day Info

#### 3.2.1. import data from scanday spreadsheet with all the experimenter information 

```{r, message = FALSE}

# spreadsheet
df_scanInfo <- read.csv(file.path(path, "raw/rawdata_scandayInfo.csv"), header = T) 

# gross clean
colnames(df_scanInfo)[1] <- "SubID"
df_scanInfo$SubID <- gsub("ML", "", df_scanInfo$SubID) 
df_scanInfo$SubID <- as.numeric(df_scanInfo$SubID) # remove ML and make numeric
df_scanInfo <- df_scanInfo[order(df_scanInfo$SubID),] # reorder according to SubID

# dplyr::select variables of interest 
runOrder <- df_scanInfo[ , grepl("Task..1|Task..2", names(df_scanInfo))]
studyPartnerName <- df_scanInfo$Study.Partner
PSI <- df_scanInfo$Slightly.Intense.PSI
anythingElseScan <- df_scanInfo$Anything.to.Note.About.the.Scan.

# Extract columns for Task 1 and Task 2 pain ratings --> state empathy pain 
painRating1 <- df_scanInfo[, grepl("Task..1.Pain.Rating", names(df_scanInfo))]
painRating2 <- df_scanInfo[, grepl("Task..2.Pain.Rating", names(df_scanInfo))]

# Initialize new columns for self and other pain ratings
df_scanInfo$painRatingSelf <- NA
df_scanInfo$painRatingOther <- NA

# Iterate over each row to assign self and other pain ratings based on run1group
for (i in 1:nrow(df_scanInfo)) {
  if (df_scanInfo$Task..1[i] == "Participant") {
    df_scanInfo$painRatingSelf[i] <- painRating1[i]
    df_scanInfo$painRatingOther[i] <- painRating2[i]
  } else if (df_scanInfo$Task..1[i] == "Study Partner") {
    df_scanInfo$painRatingSelf[i] <- painRating2[i]
    df_scanInfo$painRatingOther[i] <- painRating1[i]
  }
}

# View the updated dataframe
head(df_scanInfo)


df_scanInfo <- cbind(df_scanInfo$SubID, runOrder, studyPartnerName, PSI, anythingElseScan, painRating1, painRating2, df_scanInfo$painRatingSelf, df_scanInfo$painRatingOther)

colnames(df_scanInfo) <- c("SubID", "run1group", "run1eprime", "run2group", "run2eprime", "run1empathy", "run2empathy", "studyPartnerName", "PSI", "anythingElseScan", "painRating1", "painRating2", "painRatingSelf", "painRatingOther")

# calculate state empathy for pain, which was assessed in the scanner after each run 
df_scanInfo$stateEmp_painRating <- abs(as.numeric(df_scanInfo$painRatingSelf) - as.numeric(df_scanInfo$painRatingOther))

df_scanInfo$stateEmp_painRating <- df_scanInfo$stateEmp_painRating * (-1)

# write to csv
write.csv(df_scanInfo, file.path(path, "clean/cleandata_scandayInfo.csv")) 

```

### 3.3. Meditation Info

#### 3.3.1. import data sheet with the meditation group assignments 

```{r, message = FALSE}

# Load and clean group assignment file 
df_medGroup <- read.csv(file.path(path, "raw/meditationGroupAssignment.csv"), header = T) 
names(df_medGroup)[names(df_medGroup) == "subnum"] <- "SubID"
names(df_medGroup)[names(df_medGroup) == "condition"] <- "GroupID"
df_medGroup$SubID <- gsub("ML", "", df_medGroup$SubID) 
df_medGroup$SubID <- as.numeric(df_medGroup$SubID) # remove ML and make numeric
df_medGroup <- df_medGroup[order(df_medGroup$SubID),] # reorder according to SubID

# from other spreadsheet we know that pp 139 is GroupID LKM <- 1, so let's change this to the correct assignment 
# Replace the value in the 'GroupID' column where 'SubID' is 139 with 2
df_medGroup$GroupID[df_medGroup$SubID == 139] <- 1

df_medGroup$GroupID

```

#### 3.3.2. import data sheet with the meditation group assignments from the meditation tracking sheet

```{r, message = FALSE}

# Load and clean meditation tracking also with GroupID assignment 
df_medTracking <- read.csv(file.path(path, "raw/meditationTracking_sheet1.csv"), header = T) 

# Remove rows where any column contains "Sent" or "Completed"
df_medTracking <- df_medTracking[!apply(df_medTracking, 1, function(row) any(row %in% c("Sent", "Completed"))), ] 
df_medTracking <- df_medTracking[df_medTracking$X != "", ] # Remove rows where X column is empty
df_medTracking <- df_medTracking[, 1:2]  
colnames(df_medTracking) <- c("SubID", "GroupID") # Select only the first two columns and rename them

# Replace "LKM" with 1 and "PMR" with 2 in the group column
df_medTracking$GroupID <- ifelse(df_medTracking$GroupID == "LKM", 1, 2) 

# Remove the "ML" prefix in SubID and convert to numeric
df_medTracking$SubID <- as.numeric(gsub("ML", "", df_medTracking$SubID)) 

# View the cleaned data
head(df_medTracking)

df_medTracking$GroupID

```

#### 3.3.3. cross-check them against each other 

```{r, message = FALSE}

# Find discrepancies between the two dataframes based on SubID and GroupID
discrepancies <- merge(df_medGroup, df_medTracking, by = "SubID", suffixes = c("_MC", "_medtracking"))
discrepancies <- discrepancies[discrepancies$GroupID_MC != discrepancies$GroupID_medtracking, ]

# Print the discrepancies
if (nrow(discrepancies) > 0) {
  print(discrepancies[, c("SubID", "GroupID_MC", "GroupID_medtracking")])
} else {
  print("No discrepancies found.")
}

# Identify discrepancies between df_medGroupID and df_medTracking
discrepancies <- merge(df_medGroup, df_medTracking, by = "SubID", suffixes = c("_MC", "_medtracking"))
discrepancies <- discrepancies[discrepancies$GroupID_MC != discrepancies$GroupID_medtracking, ]

# use the ones that were in meditation tracking to replace the incorrect ones in meditation GroupID assignment 
df_medGroup$GroupID[df_medGroup$SubID == 121] <- 1
df_medGroup$GroupID[df_medGroup$SubID == 141] <- 2

# View the updated df_medGroup
head(df_medGroup)

# give meaningful labels
df_medGroup$GroupID[df_medGroup$GroupID == 1] = "LKM"
df_medGroup$GroupID[df_medGroup$GroupID == 2] = "PMR"

```

### 3.3.4. meditation tracking data frame 

```{r, message = FALSE}

# medFinish: was the meditation finished?
# medStart: when did the pp start the meditation?
# medEnd: when did the pp end the meditation?
# medNumCompleted: how many meditations were completed?
# medPercentageCompleted: what percentage of meditations were completed?

# Import meditation tracking again 
df_medTracking <- read.csv(file.path(path, "raw/meditationTracking_sheet2.csv"), header = TRUE)


# Change column names
colnames(df_medTracking) <- c("SubID", "medFinish", "medStart", "medEnd", "medNumCompleted", "medPercentageCompleted")

# Clean 'SubID' column
df_medTracking$SubID <- gsub("ML", "", df_medTracking$SubID)
df_medTracking$SubID <- as.numeric(df_medTracking$SubID)

# Clean 'medPercentageCompleted' column
df_medTracking$medPercentageCompleted <- gsub("%", "", df_medTracking$medPercentageCompleted)  # Remove percentage signs
df_medTracking$medPercentageCompleted[df_medTracking$medPercentageCompleted == "?"] <- NA  # Convert '?' to NA
df_medTracking$medPercentageCompleted <- as.numeric(df_medTracking$medPercentageCompleted)  # Convert to numeric

# Convert 'medNumCompleted' column to numeric
df_medTracking$medNumCompleted <- as.numeric(as.character(df_medTracking$medNumCompleted))

# Check the cleaned data
print(df_medTracking$medNumCompleted)
print(df_medTracking$medPercentageCompleted)


```

### 3.4. Medication Info

Medication people took on scan day 
```{r, message = FALSE}

# import data
df_medicationInfo <- read.csv(file.path(path, "raw/rawdata_medication.csv"), header = TRUE)

df_medicationInfo <- df_medicationInfo[, -((ncol(df_medicationInfo)-3):ncol(df_medicationInfo))]

# change column names: 
colnames(df_medicationInfo) <- c("SubID", "medicationsScanday")

# SubID clean
df_medicationInfo$SubID <- gsub("ML", "", df_medicationInfo$SubID) 

# make numeric
df_medicationInfo$SubID <- as.numeric(df_medicationInfo$SubID)


```


# 4. 6 months Survey (T3)

### 4.1. import data from the 6-months follow-up survey

```{r, message = FALSE}
# import data from the 6 months follow-up survey 
df_T3 <- read.csv(file.path(path, "raw/rawdata_T3_6months_postIntervention.csv"), header = TRUE)

```

### 4.2. gross clean and initial exclusions

```{r, message = FALSE}
# gross clean
df_T3 <- df_T3[grepl("ML", df_T3$id),] # exclude empty id rows
df_T3 <- df_T3[grepl(1, df_T3$Finished),] # exclude unfinished surveys
colnames(df_T3) <- gsub("id", "SubID", colnames(df_T3))
df_T3 <- df_T3 %>%
  dplyr::select(SubID, everything()) # move SubID to first row
df_T3$SubID <- gsub("ML", "", df_T3$SubID) 
df_T3$SubID <- as.numeric(df_T3$SubID) # remove ML prefix and make numeric
df_T3 <- df_T3[order(df_T3$SubID),] # reorder according to SubID

```

### 4.3. select variables of interest

```{r, message = FALSE}

# SCS
socialConnectedness <- df_T3 %>% dplyr::select(SubID, contains("SCS2"), -matches("_T_"))
reverse_cols <- c("SCS2_1", "SCS2_2", "SCS2_3", "SCS2_4","SCS2_5", "SCS2_6", "SCS2_8","SCS2_10","SCS2_15", "SCS2_18")
socialConnectedness <- data.frame(apply(socialConnectedness, 2, as.numeric))
socialConnectedness[,reverse_cols] <- 7 - socialConnectedness[,reverse_cols] # reverse code
df_long_socialConnectedness <- melt(setDT(socialConnectedness), id.vars = "SubID") # long format
df_long_socialConnectedness <- df_long_socialConnectedness %>%
  group_by(SubID)%>%
  summarise_at(vars(value), list(socialConnectedness_mean_T3 = mean)) # calculate mean per participant

df_ml_T3 <- df_long_socialConnectedness

# UCLA
loneliness <- df_T3 %>% dplyr::select(SubID, contains("UCLA2"), -matches("_T_"))
reverse_cols <- c("UCLA2_1", "UCLA2_5", "UCLA2_9", "UCLA2_10","UCLA2_15", "UCLA2_16", "UCLA2_19","UCLA2_20")
loneliness <- data.frame(apply(loneliness, 2, as.numeric))
loneliness[, reverse_cols] <- 5 - loneliness[, reverse_cols] # reverse code
df_long_loneliness <- melt(setDT(loneliness), id.vars = "SubID") # long format
df_long_loneliness <- df_long_loneliness %>%
  group_by(SubID)%>%
  summarise_at(vars(value), list(loneliness_mean_T3 = mean)) # calculate mean per participant

df_ml_T3 <- left_join(df_ml_T3, df_long_loneliness, by = "SubID")

# IRI
traitEmpathyIRI <- df_T3 %>% dplyr::select(SubID, contains("IRI2"), -matches("_T_"))
df_long_traitEmpathyIRI <- melt(setDT(traitEmpathyIRI), id.vars = "SubID") # long format
df_long_traitEmpathyIRI$value = as.numeric(df_long_traitEmpathyIRI$value) # numeric
df_long_traitEmpathyIRI <- df_long_traitEmpathyIRI %>%
  group_by(SubID)%>%
  summarise_at(vars(value), list(traitEmpathyIRI_mean_T3 = mean)) # calculate mean per participant

df_ml_T3 <- left_join(df_ml_T3, df_long_traitEmpathyIRI, by = "SubID")

#IRI subscales
#Fantasy
df_traitEmpathyIRI_fant <- traitEmpathyIRI %>% dplyr::select("SubID","IRI2_1","IRI2_5", "IRI2_7", "IRI2_12", "IRI2_16", "IRI2_23", "IRI2_26")
df_long_traitEmpathyIRI_fant <- melt(setDT(df_traitEmpathyIRI_fant), id.vars = "SubID") # long format
df_long_traitEmpathyIRI_fant$value = as.numeric(df_long_traitEmpathyIRI_fant$value)
df_long_traitEmpathyIRI_fant <- df_long_traitEmpathyIRI_fant %>%
  group_by(SubID)%>%
  summarise_at(vars(value), list(traitEmpathyIRI_fant_mean_T3 = mean)) # calculate mean per participant

df_ml_T3 <- left_join(df_ml_T3, df_long_traitEmpathyIRI_fant, by = "SubID")

#Perspective-taking
df_traitEmpathyIRI_PT <- traitEmpathyIRI %>% dplyr::select("SubID","IRI2_3","IRI2_8", "IRI2_11", "IRI2_15", "IRI2_21", "IRI2_25", "IRI2_28")
df_long_traitEmpathyIRI_PT <- melt(setDT(df_traitEmpathyIRI_PT), id.vars = "SubID") # long format
df_long_traitEmpathyIRI_PT$value = as.numeric(df_long_traitEmpathyIRI_PT$value)
df_long_traitEmpathyIRI_PT <- df_long_traitEmpathyIRI_PT %>%
  group_by(SubID)%>%
  summarise_at(vars(value), list(traitEmpathyIRI_PT_mean_T3 = mean)) # calculate mean per participant

df_ml_T3 <- left_join(df_ml_T3, df_long_traitEmpathyIRI_PT, by = "SubID")

#Empathic Concern
df_traitEmpathyIRI_EC <- traitEmpathyIRI %>% dplyr::select("SubID","IRI2_2","IRI2_4", "IRI2_9", "IRI2_14", "IRI2_18", "IRI2_20", "IRI2_22")
df_long_traitEmpathyIRI_EC <- melt(setDT(df_traitEmpathyIRI_EC), id.vars = "SubID") # long format
df_long_traitEmpathyIRI_EC$value = as.numeric(df_long_traitEmpathyIRI_EC$value)
df_long_traitEmpathyIRI_EC <- df_long_traitEmpathyIRI_EC %>%
  group_by(SubID)%>%
  summarise_at(vars(value), list(traitEmpathyIRI_EC_mean_T3 = mean)) # calculate mean per participant

df_ml_T3 <- left_join(df_ml_T3, df_long_traitEmpathyIRI_EC, by = "SubID")

#Personal Distress
df_traitEmpathyIRI_PD <- traitEmpathyIRI %>% dplyr::select("SubID","IRI2_6","IRI2_10", "IRI2_13", "IRI2_17", "IRI2_19", "IRI2_24", "IRI2_27")
df_long_traitEmpathyIRI_PD <- melt(setDT(df_traitEmpathyIRI_PD), id.vars = "SubID") # long format
df_long_traitEmpathyIRI_PD$value = as.numeric(df_long_traitEmpathyIRI_PD$value)
df_long_traitEmpathyIRI_PD <- df_long_traitEmpathyIRI_PD %>%
  group_by(SubID)%>%
  summarise_at(vars(value), list(traitEmpathyIRI_PD_mean_T3 = mean)) # calculate mean per participant

df_ml_T3 <- left_join(df_ml_T3, df_long_traitEmpathyIRI_PD, by = "SubID")

# LKC
lovingKindnessCompassion <- df_T3 %>% dplyr::select(SubID, contains("LKC2"), -matches("_T_"))
reverse_cols <- c("LKC2_3", "LKC2_4", "LKC2_7", "LKC2_11","LKC2_13", "LKC2_15")
lovingKindnessCompassion <- data.frame(apply(lovingKindnessCompassion, 2, as.numeric))
lovingKindnessCompassion[, reverse_cols] <- 5 - lovingKindnessCompassion[, reverse_cols] # reverse code
df_long_lovingKindnessCompassion <- melt(setDT(lovingKindnessCompassion), id.vars = "SubID") # long format
df_long_lovingKindnessCompassion <- df_long_lovingKindnessCompassion %>%
  group_by(SubID)%>%
  summarise_at(vars(value), list(lovingKindnessCompassion_mean_T3 = mean)) # calculate mean per participant

df_ml_T3 <- left_join(df_ml_T3, df_long_lovingKindnessCompassion, by = "SubID")

# IOS
inclusionOtherSelf <- df_T3 %>% dplyr::select(SubID, contains("IOS2"))
inclusionOtherSelf$IOS2_1b <- as.numeric(inclusionOtherSelf$IOS2_1b)
inclusionOtherSelf$IOS2_2b <- as.numeric(inclusionOtherSelf$IOS2_2b)
inclusionOtherSelf$IOS2_3b <- as.numeric(inclusionOtherSelf$IOS2_3b)
inclusionOtherSelf$IOS2_4b <- as.numeric(inclusionOtherSelf$IOS2_4b)
inclusionOtherSelf$IOS2_5 <- as.numeric(inclusionOtherSelf$IOS2_5)
inclusionOtherSelf$inclusionOtherSelf_mean <- rowMeans(inclusionOtherSelf[,c("IOS2_1b", "IOS2_2b", "IOS2_3b", "IOS2_4b", "IOS2_5")])
inclusionOtherSelf$inclusionOtherSelf_mean_nostranger <- rowMeans(inclusionOtherSelf[,c("IOS2_1b", "IOS2_2b", "IOS2_3b", "IOS2_4b")])
colnames(inclusionOtherSelf) <- c("SubID", "IOS_immediateFamily_T3", "IOS_immediateFamilyNum_T3", "IOS_extendedFamily_T3", "IOS_extendedFamilyNum_T3", "IOS_friend_T3", "IOS_friendNum_T3", "IOS_acquaintance_T3", "IOS_acquaintanceNum_T3", "IOS_strangerNum_T3", "inclusionOtherSelf_mean_T3", "inclusionOtherSelf_mean_nostranger_T3")

df_ml_T3 <- left_join(df_ml_T3, inclusionOtherSelf, by = "SubID")

# Cantrill wellbeing Ladder
wellbeing <- df_T3 %>% dplyr::select(SubID, contains("LADDER"))
wellbeing$LADDER_PRESENT2 <- as.numeric(as.character(wellbeing$LADDER_PRESENT2))
wellbeing$LADDER_FUTURE2  <- as.numeric(as.character(wellbeing$LADDER_FUTURE2))
wellbeing$wellbeing_mean_T3 <- rowMeans(wellbeing[, c("LADDER_PRESENT2", "LADDER_FUTURE2")], na.rm = TRUE)
colnames(wellbeing) <- c("SubID", "wellbeingPresent_T3", "wellbeingFuture_T3", "wellbeing_mean_T3")

df_ml_T3 <- left_join(df_ml_T3, wellbeing, by = "SubID")

```

### 4.5. post meditation experience

```{r, message = FALSE}

meditations <- df_T3 %>% dplyr::select(SubID, contains("meditations"), contains("Q422"), contains("Q423"), contains("Q424"), contains("Q426"), contains("Q427"), contains("Q429"))
#Over the past six months, have you listened to the meditations from the training (i.e., after the training was over, have you continued to listen to the meditations)?
# How many hours total did you listen to the meditations from the training in the past six months (i.e., after the training was over)?
#Over the past six months, have you practiced any forms of meditation besides the meditations from the meditation training (i.e., have you done any other type of meditation after the meditation training was over)?
#Please describe your meditation practice over the past six months (i.e., after the meditation training was over). What kind of meditation practice have you been doing and for how long?
#Over the past six months, have you had any major life change or major life event occur? 
#Please describe your major life change or event.
#Is there anything else that you would like to tell us about the past six months?

colnames(meditations) <- c("SubID", "postMedExpClosed", "postMedExpHours", "postMedExpClosedOther", "postMedExpOpen", "majorLifeChange", "majorLifeChangeDescribe", "anythingPast6months" )

df_ml_T3 <- left_join(df_ml_T3, meditations, by = "SubID")

```

### 4.6. exclusion variables and saving the file as csv

```{r, message = FALSE}

# control questions
attention <- df_T3 %>% dplyr::select(SubID, contains("ATTN2_1"), # in SCS (strongly agree - 6)
                                                contains("ATTN2_2"), # in IRI (does not describe me well - A)
                                                contains("ATTN2_3"), # in UCLA (rarely - 2)
                                                contains("Q421")) # # Do you have any questions or comments about the study? Feel free to share anything else you would like us to know.
colnames(attention) <- c("SubID", "attentionCheck1_T3", "attentionCheck2_T3", "attentionCheck3_T3", "anythingElse")

df_ml_T3 <- left_join(df_ml_T3, attention, by = "SubID")

# save
write.csv(df_ml_T3, file.path(path, "clean/cleandata_T3.csv"))

```

# 5. MVPA Results

We load the extracted ROI stats from the 3dMVM analysis in AFNI
```{r, message = FALSE}

# Read in CSV file. Adjust file path/filename if needed
df_3dmvm <- read.csv(file.path(path, "raw/3dMVM_subjectValues.csv"), header = TRUE, stringsAsFactors = FALSE)

# Rename SubjectID → SubID
names(df_3dmvm)[names(df_3dmvm) == "SubjectID"] <- "SubID"

# Remove the "sub-" prefix and leading zeros
df_3dmvm$SubID <- gsub("^sub-", "", df_3dmvm$SubID)
df_3dmvm$SubID <- sub("^0+", "", df_3dmvm$SubID)

# Replace dots
colnames(df_3dmvm) <- gsub("\\.", "_", colnames(df_3dmvm))
colnames(df_3dmvm) <- gsub("X3dMVM_Group", "", colnames(df_3dmvm))
colnames(df_3dmvm) <- gsub("X3dMVM_Groupmc_", "", colnames(df_3dmvm))
colnames(df_3dmvm) <- gsub("p0_001_", "", colnames(df_3dmvm))
colnames(df_3dmvm) <- gsub("_t", "", colnames(df_3dmvm))

# Convert to numeric (optional but often helpful)
df_3dmvm$SubID <- as.numeric(df_3dmvm$SubID)

```

# 6. Merge all data frames

dfB_T1T2: 148 variables and 117 observations because purely behavioral data without 6 months or scan day behavioral data 
df_behavior: 155 variables and 108 observations bc group column is in there now 
dfB_all: 202 variables and 108 participants because we added 6 months survey to the above data frame. Not everyone did 6 months survey (only 102 pps), so we have some NAs in this df. 
df_scanSurvey1: 202 variables and 59 participants
df_scanSurvey2: 211 variables and 59 participants
df_scanSurvey3: 10 variables and 54 participants
df_scanSurvey4: 47 variables and 54 participants (now the ones who were excluded are not in there anymore)
df_scanSurvey5: 257 variables and 54 participants
dfS_all: 303 variables and 54 participants

108 not long and 216 long because we removed fmri people from behavioral sample but not from neuro sample based on attention checks

### 6.1. Exclusions

```{r, message = FALSE}

# for behavioral analysis, retain only the participants that also did the 4 weeks
dfB_T1T2 <- merge(df_ml_T1, df_ml_T2) # 149 variables and 117 participants 
dfB_noT3 <- merge(dfB_T1T2,df_medGroup) # 150 variables and 110 participants 
dfB_noT3 <- left_join(dfB_noT3, df_medTracking, by = "SubID") # 155 variables and 110 participants
dfB_noT3 <- left_join(dfB_noT3, df_medicationInfo, by = "SubID") # 156 variables and 110 participants
dfB_all <- left_join(dfB_noT3, df_ml_T3, by = "SubID") # 202 variables and 110 participants 

# exclude people that fail attention checks at least twice
calculate_majority_value = function(column) { 
  majority_value = as.numeric(names(which.max(table(column)))) # calculate the majority value for each column
  return(majority_value)
}

# List of SubIDs that should not be excluded (see participants section for details)
SubIDs_to_keep <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 14, 18, 31, 34, 39, 43, 50, 52, 53, 57, 58, 63, 64, 67, 71, 74, 75, 76, 81, 87, 88, 91, 102, 106, 109, 115, 119, 122, 130, 136, 138, 139, 149, 154, 156, 158, 159, 167, 169, 170, 179, 180, 187, 191, 192, 193, 194, 195, 196, 197)

# Calculate majority values for each attention column
attention_columns <- dfB_all %>% select(contains("attention"))
majority_values <- sapply(attention_columns, calculate_majority_value)

# Initialize a vector to store the mismatch counts
mismatch_counts <- numeric(nrow(dfB_all))

# Count mismatches for each SubID
for (i in 1:nrow(dfB_all)) {
  count <- 0
  for (colname in names(majority_values)) {
    SubID_value = as.numeric(dfB_all[i, colname])
    majority_value = majority_values[colname]
    if (!is.na(SubID_value) && SubID_value != majority_value) {
      count <- count + 1
    }
  }
  mismatch_counts[i] <- count
}

# Add the mismatch counts as a new column to the dataframe
dfB_all$attention_diff_count <- mismatch_counts

# Print the SubID where count is larger or equal than 2 and not in the list to keep
SubIDs_to_exclude <- dfB_all %>%
  filter(attention_diff_count >= 2, !SubID %in% SubIDs_to_keep) %>%
  select(SubID)

print(SubIDs_to_exclude)

SubIDs_to_exclude_also_scan <- dfB_all %>%
  filter(attention_diff_count >= 2,) %>%
  select(SubID)

print(SubIDs_to_exclude_also_scan)

# Exclude these rows from the dfB_all dataframe, while keeping the specified SubIDs because we still need to make scan frame and there we don't exclude the pps from because of power 
dfB_all <- dfB_all %>%
  filter(attention_diff_count < 2 | SubID %in% SubIDs_to_keep)

# Exclude the same rows from the dfB_noT3 dataframe
dfB_noT3 <- dfB_noT3 %>%
  filter(!(SubID %in% SubIDs_to_exclude$SubID))

```

### 6.2. save clean files

```{r, message = FALSE}

# for neuro analysis, include scanday survey, spreadsheet information, and pattern similarity & AUC scores
dfS <- merge(dfB_noT3, df_ml_scanSurvey) # add scan day data (202 variables and 59 participants)
dfS <- merge(dfS, df_scanInfo) # add scan spreadsheet (211 variables and 59 participants)
dfS_noT3 <- left_join(df_3dmvm,dfS, by = "SubID")
dfS_all <- left_join(dfS_noT3, df_ml_T3, by = "SubID") # 303 variables and 54 participants

## exclude the scan participants from the behavioral sample that didn't pass the attention check in the scan sample as seen above (SubID 76 and 191)
dfB_all <- dfB_all %>%
  filter(attention_diff_count < 2)

# Exclude the same rows from all final dfs
dfB_noT3 <- dfB_noT3 %>%
  filter(!(SubID %in% SubIDs_to_exclude_also_scan$SubID))

# write to csv
# behavior including scan data but not T3 because some people dropped out for T3 and we don't want to lose them in this df
write.csv(dfS_noT3, file.path(path, "clean/cleandata_scanSample_noT3.csv"))

# behavior including scan data 6 months for multilevel models to assess changes over time 
write.csv(dfS_all, file.path(path, "clean/cleandata_scanSample_all.csv"))

# behavior only
write.csv(dfB_noT3, file.path(path, "clean/cleandata_fullSample_noT3.csv"))

# behavior only 6 months
write.csv(dfB_all, file.path(path, "clean/cleandata_fullSample_all.csv"))

```

### 6.3 long format for MLLM: full sample 

```{r, message = FALSE}

df <- dfB_noT3

# Identify columns with T1, T2 in their names
columns_t1 <- grep("T1", names(df), value = TRUE)
columns_t2 <- grep("T2", names(df), value = TRUE)

# Print the identified columns for debugging
print(columns_t1)
print(columns_t2)

# Find all unique base column names by removing _T1, _T2, _T3
base_columns <- unique(gsub("_T[12]", "", c(columns_t1, columns_t2)))

# Print the base column names for debugging
print(base_columns)

# Initialize list to store dataframes
df_list <- list()

if (is.data.frame(df)) {
  for (base_col in base_columns) {
    col_t1 <- paste0(base_col, "_T1")
    col_t2 <- paste0(base_col, "_T2")

    # Print the column names for debugging
    print(paste("Processing:", col_t1, col_t2))
    
    # Create temporary dataframes for each time point, if they exist
    temp_t1 <- if (col_t1 %in% names(df)) {
      df %>%
        select(SubID, !!base_col := !!sym(col_t1)) %>%
        mutate(time = 1)
    } else {
      df %>%
        select(SubID) %>%
        mutate(!!base_col := NA, time = 1)
    }
    
    temp_t2 <- if (col_t2 %in% names(df)) {
      df %>%
        select(SubID, !!base_col := !!sym(col_t2)) %>%
        mutate(time = 2)
    } else {
      df %>%
        select(SubID) %>%
        mutate(!!base_col := NA, time = 2)
    }
    
    
    # Combine the temporary dataframes
    temp_df <- bind_rows(temp_t1, temp_t2)
    
    # Print the combined dataframe for debugging
    #print(temp_df)
    
    # Append to the list
    df_list[[base_col]] <- temp_df
  }
  
  # Combine all long dataframes by SubID and time
  df_long <- reduce(df_list, full_join, by = c("SubID", "time"))
  
  # Print the combined long dataframe for debugging
  print(head(df_long, 10)) # Displaying the first 10 rows for a quick check
  
  # Joining with the original dataframe
  df_behavior_long <- left_join(df_long, df, by = "SubID")
  
  # Print the final long-format dataframe for debugging
  print(head(df_behavior_long, 10))
  
  # Save the long-format dataframe
  write.csv(df_behavior_long, file.path(path, "clean/cleandata_fullSample_noT3_long.csv"))
} else {
  print("Error: df is not a data frame. Please check your dataframe definition.")
}


```

### 6.4 long format for MLM: full sample 6 months

```{r, message = FALSE}

df <- dfB_all

# Identify columns with T1, T2, and T3 in their names
columns_t1 <- grep("T1", names(df), value = TRUE)
columns_t2 <- grep("T2", names(df), value = TRUE)
columns_t3 <- grep("T3", names(df), value = TRUE)

# Print the identified columns for debugging
print(columns_t1)
print(columns_t2)
print(columns_t3)

# Find all unique base column names by removing _T1, _T2, _T3
base_columns <- unique(gsub("_T[123]", "", c(columns_t1, columns_t2, columns_t3)))

# Print the base column names for debugging
print(base_columns)

# Initialize list to store dataframes
df_list <- list()

if (is.data.frame(df)) {
  for (base_col in base_columns) {
    col_t1 <- paste0(base_col, "_T1")
    col_t2 <- paste0(base_col, "_T2")
    col_t3 <- paste0(base_col, "_T3")
    
    # Print the column names for debugging
    print(paste("Processing:", col_t1, col_t2, col_t3))
    
    # Create temporary dataframes for each time point, if they exist
    temp_t1 <- if (col_t1 %in% names(df)) {
      df %>%
        select(SubID, !!base_col := !!sym(col_t1)) %>%
        mutate(time = 1)
    } else {
      df %>%
        select(SubID) %>%
        mutate(!!base_col := NA, time = 1)
    }
    
    temp_t2 <- if (col_t2 %in% names(df)) {
      df %>%
        select(SubID, !!base_col := !!sym(col_t2)) %>%
        mutate(time = 2)
    } else {
      df %>%
        select(SubID) %>%
        mutate(!!base_col := NA, time = 2)
    }
    
    temp_t3 <- if (col_t3 %in% names(df)) {
      df %>%
        select(SubID, !!base_col := !!sym(col_t3)) %>%
        mutate(time = 3)
    } else {
      df %>%
        select(SubID) %>%
        mutate(!!base_col := NA, time = 3)
    }
    
    # Print the temporary dataframes for debugging
    print(temp_t1)
    print(temp_t2)
    print(temp_t3)
    
    # Combine the temporary dataframes
    temp_df <- bind_rows(temp_t1, temp_t2, temp_t3)
    
    # Print the combined dataframe for debugging
   # print(temp_df)
    
    # Append to the list
    df_list[[base_col]] <- temp_df
  }
  
  # Combine all long dataframes by SubID and time
  df_long <- reduce(df_list, full_join, by = c("SubID", "time"))
  
  # Print the combined long dataframe for debugging
  print(head(df_long, 10)) # Displaying the first 10 rows for a quick check
  
  # Joining with the original dataframe
  dfB_all_long <- left_join(df_long, df, by = "SubID")
  
  # Print the final long-format dataframe for debugging
  print(head(dfB_all_long, 10))
  
  # Save the long-format dataframe
  write.csv(dfB_all_long,  file.path(path, "clean/cleandata_fullSample_all_long.csv"))
} else {
  print("Error: df is not a data frame. Please check your dataframe definition.")
}

```

### 6.5 long format for MLM: scan sample without T3

```{r, message = FALSE}

# Define the dataframe
df <- dfS_noT3

# Identify columns with T1, T2, and T3 in their names
columns_t1 <- grep("T1", names(df), value = TRUE)
columns_t2 <- grep("T2", names(df), value = TRUE)

# Print the identified columns for debugging
print(columns_t1)
print(columns_t2)

# Find all unique base column names by removing _T1, _T2, _T3
base_columns <- unique(gsub("_T[12]", "", c(columns_t1, columns_t2)))

# Print the base column names for debugging
print(base_columns)

# Initialize list to store dataframes
df_list <- list()

if (is.data.frame(df)) {
  for (base_col in base_columns) {
    col_t1 <- paste0(base_col, "_T1")
    col_t2 <- paste0(base_col, "_T2")

    # Print the column names for debugging
    print(paste("Processing:", col_t1, col_t2))
    
    # Create temporary dataframes for each time point, if they exist
    temp_t1 <- if (col_t1 %in% names(df)) {
      df %>%
        select(SubID, !!base_col := !!sym(col_t1)) %>%
        mutate(time = 1)
    } else {
      df %>%
        select(SubID) %>%
        mutate(!!base_col := NA, time = 1)
    }
    
    temp_t2 <- if (col_t2 %in% names(df)) {
      df %>%
        select(SubID, !!base_col := !!sym(col_t2)) %>%
        mutate(time = 2)
    } else {
      df %>%
        select(SubID) %>%
        mutate(!!base_col := NA, time = 2)
    }
    
    
    # Combine the temporary dataframes
    temp_df <- bind_rows(temp_t1, temp_t2)
    
    # Print the combined dataframe for debugging
    #print(temp_df)
    
    # Append to the list
    df_list[[base_col]] <- temp_df
  }
  
  # Combine all long dataframes by SubID and time
  df_long <- reduce(df_list, full_join, by = c("SubID", "time"))
  
  # Print the combined long dataframe for debugging
  print(head(df_long, 10)) # Displaying the first 10 rows for a quick check
  
  # Joining with the original dataframe
  dfS_noT3_long <- left_join(df_long, df, by = "SubID")
  
  # Print the final long-format dataframe for debugging
  print(head(dfS_noT3_long, 10))
  
  # Save the long-format dataframe
  write.csv(dfS_noT3_long, file.path(path, "clean/cleandata_scanSample_noT3_long.csv"))
} else {
  print("Error: df is not a data frame. Please check your dataframe definition.")
}



```

### 6.6 long format for MLM: scan sample 6 months included

```{r, message = FALSE}

df <- dfS_all

# Identify columns with T1, T2, and T3 in their names
columns_t1 <- grep("T1", names(df), value = TRUE)
columns_t2 <- grep("T2", names(df), value = TRUE)
columns_t3 <- grep("T3", names(df), value = TRUE)

# Print the identified columns for debugging
print(columns_t1)
print(columns_t2)
print(columns_t3)

# Find all unique base column names by removing _T1, _T2, _T3
base_columns <- unique(gsub("_T[123]", "", c(columns_t1, columns_t2, columns_t3)))

# Print the base column names for debugging
print(base_columns)

# Initialize list to store dataframes
df_list <- list()

if (is.data.frame(df)) {
  for (base_col in base_columns) {
    col_t1 <- paste0(base_col, "_T1")
    col_t2 <- paste0(base_col, "_T2")
    col_t3 <- paste0(base_col, "_T3")
    
    # Print the column names for debugging
    print(paste("Processing:", col_t1, col_t2, col_t3))
    
    # Create temporary dataframes for each time point, if they exist
    temp_t1 <- if (col_t1 %in% names(df)) {
      df %>%
        select(SubID, !!base_col := !!sym(col_t1)) %>%
        mutate(time = 1)
    } else {
      df %>%
        select(SubID) %>%
        mutate(!!base_col := NA, time = 1)
    }
    
    temp_t2 <- if (col_t2 %in% names(df)) {
      df %>%
        select(SubID, !!base_col := !!sym(col_t2)) %>%
        mutate(time = 2)
    } else {
      df %>%
        select(SubID) %>%
        mutate(!!base_col := NA, time = 2)
    }
    
    temp_t3 <- if (col_t3 %in% names(df)) {
      df %>%
        select(SubID, !!base_col := !!sym(col_t3)) %>%
        mutate(time = 3)
    } else {
      df %>%
        select(SubID) %>%
        mutate(!!base_col := NA, time = 3)
    }
    
    # Print the temporary dataframes for debugging
    print(temp_t1)
    print(temp_t2)
    print(temp_t3)
    
    # Combine the temporary dataframes
    temp_df <- bind_rows(temp_t1, temp_t2, temp_t3)
    
    # Print the combined dataframe for debugging
    #print(temp_df)
    
    # Append to the list
    df_list[[base_col]] <- temp_df
  }
  
  # Combine all long dataframes by SubID and time
  df_long <- reduce(df_list, full_join, by = c("SubID", "time"))
  
  # Print the combined long dataframe for debugging
  print(head(df_long, 10)) # Displaying the first 10 rows for a quick check
  
  # Joining with the original dataframe
  dfS_all_long <- left_join(df_long, df, by = "SubID")
  
  # Print the final long-format dataframe for debugging
  print(head(dfS_all_long, 10))
  
  # Save the long-format dataframe
  write.csv(dfS_all_long,file.path(path, "clean/cleandata_scanSample_all_long.csv"))
} else {
  print("Error: df is not a data frame. Please check your dataframe definition.")
}



```

# 7. Check pp number

```{r, message = FALSE}

table(dfB_noT3$GroupID)
table(dfS_noT3$GroupID)
```
