---
title: "Loneliness modulates self-reported empathy but not empathic multi-voxel neural response patterns after meditation training"
author: "Marla Dressel"
date: "`r Sys.Date()`"
output: html_document
---

This script performs analysis for "Loneliness modulates self-reported empathy but not empathic multi-voxel neural response patterns after meditation training". See preregistration for details: https://osf.io/agbu5

# Preliminary Analysis

## 1. Setting up the script

### 1.1. Loading packages

```{r, message = FALSE}
# chronological order of usage
library(gt) # advanced tables
library(dplyr) # %>%
library(janitor) # adorn_ns
library(kableExtra)  # apa table stuff
library(ggplot2) # ggplot
library(papaja) # theme_apa
library(gridExtra) # grid.arrange
library(moments) # skewness
library(openxlsx) # write.xlsx
library(reshape2) # cor table
library(nlme) # lme function
library(sjPlot) # tab_model
library(lmtest) #lrtest
```

### 1.2. Importing the full clean data

```{r, message = FALSE}

# for T2 analysis, we use a wide df. 


path <- "/your/path/Behavioral_Data_R_Scripts/data" # CHANGE THIS PATH TO YOUR DIRECTORY!

dfB_noT3 <- read.csv(file.path(path, "clean/cleandata_fullSample_noT3.csv"), row.names = 1)
dfB_noT3_long <- read.csv(file.path(path, "clean/cleandata_fullSample_noT3_long.csv"), row.names = 1)
dfB_all_long <- read.csv(file.path(path, "clean/cleandata_fullSample_all_long.csv"), row.names = 1) # for analysis over time 

# Ensure the group variable is a factor and PMR is the reference group (control)
dfB_noT3$GroupID <- as.factor(dfB_noT3$GroupID)
dfB_noT3$GroupID <- relevel(dfB_noT3$GroupID, ref = "PMR")
dfB_noT3_long$GroupID <- as.factor(dfB_noT3_long$GroupID)
dfB_noT3_long$GroupID <- relevel(dfB_noT3_long$GroupID, ref = "PMR")
dfB_all_long$GroupID <- as.factor(dfB_all_long$GroupID)
dfB_all_long$GroupID <- relevel(dfB_all_long$GroupID, ref = "PMR")

```

### 1.3. Importing the scan sample 

```{r, message = FALSE}

dfS_noT3 <- read.csv(file.path(path, "clean/cleandata_scanSample_noT3.csv"), row.names = 1)
dfS_noT3_long <- read.csv(file.path(path, "clean/cleandata_scanSample_noT3_long.csv"), row.names = 1)
dfS_noT3_all_long <- read.csv(file.path(path, "clean/cleandata_scanSample_all_long.csv"), row.names = 1)

# Ensure the group variable is a factor and PMR is the reference group (control)
dfS_noT3$GroupID <- as.factor(dfS_noT3$GroupID)
dfS_noT3$GroupID <- relevel(dfS_noT3$GroupID, ref = "PMR")
dfS_noT3_long$GroupID <- as.factor(dfS_noT3_long$GroupID)
dfS_noT3_long$GroupID <- relevel(dfS_noT3_long$GroupID, ref = "PMR")
dfS_noT3_all_long$GroupID <- as.factor(dfS_noT3_all_long$GroupID)
dfS_noT3_all_long$GroupID <- relevel(dfS_noT3_all_long$GroupID, ref = "PMR")

```

## 2. Demographics

### 2.1. Differences in age between samples

```{r, message = FALSE}

library(rstatix)
# Create the data frame
df_age <- as.data.frame(cbind(dfB_noT3$age, dfS_noT3$age))
colnames(df_age) <- c("dB1", "dB2")

# Calculate means and standard deviations
mean_dB1 <- mean(df_age$dB1)
mean_dB2 <- mean(df_age$dB2)
sd_dB1 <- sd(df_age$dB1)
sd_dB2 <- sd(df_age$dB2)

# Calculate the standard errors
se_dB1 <- sd_dB1 / sqrt(length(df_age$dB1))
se_dB2 <- sd_dB2 / sqrt(length(df_age$dB2))

# Calculate the 95% confidence intervals
alpha <- 0.05
t_critical <- qt(1 - alpha / 2, df = length(df_age$dB1) - 1)

ci_dB1_lower <- mean_dB1 - t_critical * se_dB1
ci_dB1_upper <- mean_dB1 + t_critical * se_dB1

ci_dB2_lower <- mean_dB2 - t_critical * se_dB2
ci_dB2_upper <- mean_dB2 + t_critical * se_dB2

# Perform the t-test
t_test <- t.test(df_age$dB1, df_age$dB2)

# Calculate Cohen's d and its confidence interval using compute.es
if (!require(compute.es)) {
  install.packages("compute.es")
  library(compute.es)
}

n1 <- length(df_age$dB1)
n2 <- length(df_age$dB2)
cohens_d <- (mean_dB1 - mean_dB2) / sqrt(((n1 - 1) * sd_dB1^2 + (n2 - 1) * sd_dB2^2) / (n1 + n2 - 2))

es <- tes(t_test$statistic, n1, n2, dig = 3)

# Print the results
cat("Mean of dB1:", mean_dB1, "\n")
cat("95% CI for mean of dB1:", ci_dB1_lower, "to", ci_dB1_upper, "\n")
cat("Mean of dB2:", mean_dB2, "\n")
cat("95% CI for mean of dB2:", ci_dB2_lower, "to", ci_dB2_upper, "\n")
print(t_test)
cat("Cohen's d:", cohens_d, "\n")
cat("95% CI for Cohen's d:", es$l.d, "to", es$u.d, "\n")

```

### 2.1. Differences in age between samples

```{r, message = FALSE}

library(rstatix)
# Create the data frame
df_age <- as.data.frame(cbind(dfB_noT3$loneliness_mean_T1, dfS_noT3$loneliness_mean_T1))
colnames(df_age) <- c("dB1", "dB2")

# Calculate means and standard deviations
mean_dB1 <- mean(df_age$dB1)
mean_dB2 <- mean(df_age$dB2)
sd_dB1 <- sd(df_age$dB1)
sd_dB2 <- sd(df_age$dB2)

# Calculate the standard errors
se_dB1 <- sd_dB1 / sqrt(length(df_age$dB1))
se_dB2 <- sd_dB2 / sqrt(length(df_age$dB2))

# Calculate the 95% confidence intervals
alpha <- 0.05
t_critical <- qt(1 - alpha / 2, df = length(df_age$dB1) - 1)

ci_dB1_lower <- mean_dB1 - t_critical * se_dB1
ci_dB1_upper <- mean_dB1 + t_critical * se_dB1

ci_dB2_lower <- mean_dB2 - t_critical * se_dB2
ci_dB2_upper <- mean_dB2 + t_critical * se_dB2

# Perform the t-test
t_test <- t.test(df_age$dB1, df_age$dB2)

# Calculate Cohen's d and its confidence interval using compute.es
if (!require(compute.es)) {
  install.packages("compute.es")
  library(compute.es)
}

n1 <- length(df_age$dB1)
n2 <- length(df_age$dB2)
cohens_d <- (mean_dB1 - mean_dB2) / sqrt(((n1 - 1) * sd_dB1^2 + (n2 - 1) * sd_dB2^2) / (n1 + n2 - 2))

es <- tes(t_test$statistic, n1, n2, dig = 3)

# Print the results
cat("Mean of dB1:", mean_dB1, "\n")
cat("95% CI for mean of dB1:", ci_dB1_lower, "to", ci_dB1_upper, "\n")
cat("Mean of dB2:", mean_dB2, "\n")
cat("95% CI for mean of dB2:", ci_dB2_lower, "to", ci_dB2_upper, "\n")
print(t_test)
cat("Cohen's d:", cohens_d, "\n")
cat("95% CI for Cohen's d:", es$l.d, "to", es$u.d, "\n")

```

### 2.2. Calculate frequencies and percentages for categorical demographic variables.  

##### 2.2.1. Full Sample

```{r, message = FALSE}

library(janitor) # adorn_ns
# Generating frequency table for age
age_table <- dfB_noT3 %>%
  tabyl(agefull, GroupID) %>%
  adorn_totals("row") %>% # Adds total counts per age
  adorn_totals("col") %>% # Adds total counts per GroupID at the bottom
  adorn_percentages("col") %>% # Converts counts to percentages by column (GroupID)
  adorn_pct_formatting(digits = 1) %>%
  adorn_ns(position = "front")

# Generating frequency table for Gender
gender_table <- dfB_noT3 %>%
  tabyl(gender, GroupID) %>%
  adorn_totals("row") %>% # Adds total counts per gender
  adorn_totals("col") %>% # Adds total counts per GroupID at the bottom
  adorn_percentages("col") %>% # Converts counts to percentages by column (GroupID)
  adorn_pct_formatting(digits = 1) %>%
  adorn_ns(position = "front")

# Generating frequency table for Ethnicity
ethnicity_table <- dfB_noT3 %>%
  tabyl(ethnicity, GroupID) %>%
  adorn_totals("row") %>% # Adds total counts per ethnicity
  adorn_totals("col") %>% # Adds total counts per GroupID at the bottom
  adorn_percentages("col") %>% # Converts counts to percentages by column (GroupID)
  adorn_pct_formatting(digits = 1) %>%
  adorn_ns(position = "front")

# Frequency table for Household Income by GroupID
income_table <- dfB_noT3 %>%
  tabyl(income, GroupID) %>%
  adorn_totals("row") %>% # Adds total counts per income
  adorn_totals("col") %>% # Adds total counts per GroupID at the bottom
  adorn_percentages("col") %>% # Converts counts to percentages by column (GroupID)
  adorn_pct_formatting(digits = 1) %>%
  adorn_ns(position = "front")

# Frequency table for Education by GroupID
education_table <- dfB_noT3 %>%
  tabyl(education, GroupID) %>%
  adorn_totals("row") %>% # Adds total counts per education
  adorn_totals("col") %>% # Adds total counts per GroupID at the bottom
  adorn_percentages("col") %>% # Converts counts to percentages by column (GroupID)
  adorn_pct_formatting(digits = 1) %>%
  adorn_ns(position = "front")

# Frequency table for Education by GroupID
employment_table <- dfB_noT3 %>%
  tabyl(employment, GroupID) %>%
  adorn_totals("row") %>% # Adds total counts per employment 
  adorn_totals("col") %>% # Adds total counts per GroupID at the bottom
  adorn_percentages("col") %>% # Converts counts to percentages by column (GroupID)
  adorn_pct_formatting(digits = 1) %>%
  adorn_ns(position = "front")

# Frequency table for handedness by GroupID
handedness_table <- dfB_noT3 %>%
  tabyl(handedness, GroupID) %>%
  adorn_totals("row") %>% # Adds total counts per handedness
  adorn_totals("col") %>% # Adds total counts per GroupID at the bottom
  adorn_percentages("col") %>% # Converts counts to percentages by column (GroupID)
  adorn_pct_formatting(digits = 1) %>%
  adorn_ns(position = "front")

print(age_table)
print(gender_table)
print(ethnicity_table)
print(education_table)
print(employment_table)
print(income_table)
print(handedness_table)

```

##### 2.2.2. Scan Sample

```{r, message = FALSE}
###### scan!

# Generating frequency table for age
age_table_scan <- dfS_noT3 %>%
  tabyl(agefull, GroupID) %>%
  adorn_totals("row") %>% # Adds total counts per age
  adorn_totals("col") %>% # Adds total counts per GroupID at the bottom
  adorn_percentages("col") %>% # Converts counts to percentages by column (GroupID)
  adorn_pct_formatting(digits = 1) %>%
  adorn_ns(position = "front")

# Generating frequency table for Gender
gender_table_scan <- dfS_noT3 %>%
  tabyl(gender, GroupID) %>%
  adorn_totals("row") %>% # Adds total counts per gender
  adorn_totals("col") %>% # Adds total counts per GroupID at the bottom
  adorn_percentages("col") %>% # Converts counts to percentages by column (GroupID)
  adorn_pct_formatting(digits = 1) %>%
  adorn_ns(position = "front")

# Generating frequency table for Ethnicity
ethnicity_table_scan <- dfS_noT3 %>%
  tabyl(ethnicity, GroupID) %>%
  adorn_totals("row") %>% # Adds total counts per ethnicity 
  adorn_totals("col") %>% # Adds total counts per GroupID at the bottom
  adorn_percentages("col") %>% # Converts counts to percentages by column (GroupID)
  adorn_pct_formatting(digits = 1) %>%
  adorn_ns(position = "front")

# Frequency table for Household Income by GroupID
income_table_scan <- dfS_noT3 %>%
  tabyl(income, GroupID) %>%
  adorn_totals("row") %>% # Adds total counts per income
  adorn_totals("col") %>% # Adds total counts per GroupID at the bottom
  adorn_percentages("col") %>% # Converts counts to percentages by column (GroupID)
  adorn_pct_formatting(digits = 1) %>%
  adorn_ns(position = "front")

# Frequency table for Education by GroupID
education_table_scan <- dfS_noT3 %>%
  tabyl(education, GroupID) %>%
  adorn_totals("row") %>% # Adds total counts per education
  adorn_totals("col") %>% # Adds total counts per GroupID at the bottom
  adorn_percentages("col") %>% # Converts counts to percentages by column (GroupID)
  adorn_pct_formatting(digits = 1) %>%
  adorn_ns(position = "front")

# Frequency table for Education by GroupID
employment_table_scan <- dfS_noT3 %>%
  tabyl(employment, GroupID) %>%
  adorn_totals("row") %>% # Adds total counts per employment 
  adorn_totals("col") %>% # Adds total counts per GroupID at the bottom
  adorn_percentages("col") %>% # Converts counts to percentages by column (GroupID)
  adorn_pct_formatting(digits = 1) %>%
  adorn_ns(position = "front")

# Frequency table for Education by GroupID
handedness_table_scan <- dfS_noT3 %>%
  tabyl(handedness, GroupID) %>%
  adorn_totals("row") %>% # Adds total counts per handedness
  adorn_totals("col") %>% # Adds total counts per GroupID at the bottom
  adorn_percentages("col") %>% # Converts counts to percentages by column (GroupID)
  adorn_pct_formatting(digits = 1) %>%
  adorn_ns(position = "front")


print(age_table_scan)
print(gender_table_scan)
print(ethnicity_table_scan)
print(education_table_scan)
print(employment_table_scan)
print(income_table_scan)
print(handedness_table_scan)

```

### 2.3. Test whether there are significant differences within each demographic category (full Sample vs scan sample).

```{r, message = FALSE}
# Function to perform proportion test and print results
perform_proportion_test <- function(counts) {
  prop_test <- prop.test(counts)
  test_statistic <- prop_test$statistic
  p_value <- prop_test$p.value
  prop1 <- counts[1, 1] / sum(counts[1, ])
  prop2 <- counts[2, 1] / sum(counts[2, ])
  effect_size <- 2 * asin(sqrt(prop1)) - 2 * asin(sqrt(prop2))
  
  # Calculate the confidence interval for the difference in proportions
  ci_diff <- prop_test$conf.int
  ci_effect_size_lower <- 2 * asin(sqrt((ci_diff[1] + 1) / 2)) - 2 * asin(sqrt((1 - ci_diff[1]) / 2))
  ci_effect_size_upper <- 2 * asin(sqrt((ci_diff[2] + 1) / 2)) - 2 * asin(sqrt((1 - ci_diff[2]) / 2))
  
  cat("Proportion in dfB_noT3:", prop1, "\n")
  cat("Proportion in dfS_noT3:", prop2, "\n")
  cat("Test Statistic:", test_statistic, "\n")
  cat("P-value:", p_value, "\n")
  cat("Effect Size (Cohen's h):", effect_size, "\n")
  cat("95% CI for Effect Size (Cohen's h):", ci_effect_size_lower, ci_effect_size_upper, "\n")
}

# Gender: Female vs Male
cat("\nTesting Gender: Female vs Male\n")
counts_gender <- matrix(c(sum(dfB_noT3$gender == "female"), sum(dfB_noT3$gender == "male"),
                          sum(dfS_noT3$gender == "female"), sum(dfS_noT3$gender == "male")), nrow = 2)
perform_proportion_test(counts_gender)

# Ethnicity: White vs all other ethnicities
cat("\nTesting Ethnicity: White vs all other ethnicities\n")
counts_ethnicity <- matrix(c(sum(dfB_noT3$ethnicity == "White"), sum(dfB_noT3$ethnicity != "White"),
                             sum(dfS_noT3$ethnicity == "White"), sum(dfS_noT3$ethnicity != "White")), nrow = 2)
perform_proportion_test(counts_ethnicity)

# Ethnicity: Black vs all other ethnicities
cat("\nTesting Ethnicity: Black vs all other ethnicities\n")
counts_black <- matrix(c(sum(dfB_noT3$ethnicity == "Black"), sum(dfB_noT3$ethnicity != "Black"),
                         sum(dfS_noT3$ethnicity == "Black"), sum(dfS_noT3$ethnicity != "Black")), nrow = 2)
perform_proportion_test(counts_black)

# Ethnicity: Asian vs all other ethnicities
cat("\nTesting Ethnicity: Asian vs all other ethnicities\n")
counts_asian <- matrix(c(sum(dfB_noT3$ethnicity == "Asian"), sum(dfB_noT3$ethnicity != "Asian"),
                         sum(dfS_noT3$ethnicity == "Asian"), sum(dfS_noT3$ethnicity != "Asian")), nrow = 2)
perform_proportion_test(counts_asian)

# Ethnicity: Hispanic vs all other ethnicities
cat("\nTesting Ethnicity: Hispanic vs all other ethnicities\n")
counts_hispanic <- matrix(c(sum(dfB_noT3$ethnicity == "Hispanic"), sum(dfB_noT3$ethnicity != "Hispanic"),
                            sum(dfS_noT3$ethnicity == "Hispanic"), sum(dfS_noT3$ethnicity != "Hispanic")), nrow = 2)
perform_proportion_test(counts_hispanic)

# Ethnicity: Other vs all other ethnicities
cat("\nTesting Ethnicity: Other vs all other ethnicities\n")
counts_other <- matrix(c(sum(dfB_noT3$ethnicity == "Other"), sum(dfB_noT3$ethnicity != "Other"),
                         sum(dfS_noT3$ethnicity == "Other"), sum(dfS_noT3$ethnicity != "Other")), nrow = 2)
perform_proportion_test(counts_other)

# Income: Upper vs Lower
cat("\nTesting Income: Upper vs Lower\n")
counts_income <- matrix(c(sum(dfB_noT3$income == "Upper"), sum(dfB_noT3$income == "Lower"),
                          sum(dfS_noT3$income == "Upper"), sum(dfS_noT3$income == "Lower")), nrow = 2)
perform_proportion_test(counts_income)

# Education: YesCollegeDegree vs NoCollegeDegree
cat("\nTesting Education: YesCollegeDegree vs NoCollegeDegree\n")
counts_education <- matrix(c(sum(dfB_noT3$education == "YesCollegeDegree"), sum(dfB_noT3$education == "NoCollegeDegree"),
                             sum(dfS_noT3$education == "YesCollegeDegree"), sum(dfS_noT3$education == "NoCollegeDegree")), nrow = 2)
perform_proportion_test(counts_education)

# Handedness: Right vs Left
cat("\nTesting Handedness: Right vs Left\n")
counts_handedness <- matrix(c(sum(dfB_noT3$handedness == "Right"), sum(dfB_noT3$handedness == "Left"),
                              sum(dfS_noT3$handedness == "Right"), sum(dfS_noT3$handedness == "Left")), nrow = 2)
perform_proportion_test(counts_handedness)

```


Primary Analysis

## 3. Pre-Intervention analysis 

### 3.1. Differences at baseline full sample 

```{r, message = FALSE}

library(effectsize)
conflicted::conflicts_prefer(effectsize::cohens_d)

perform_full_analysis <- function(variable_name, data) {
  
  lkm_data <- data[data$GroupID == "LKM", variable_name]
  pmr_data <- data[data$GroupID == "PMR", variable_name]
  
  ## welch's two-sample two-sided t-test
  t_test <- t.test(lkm_data, pmr_data)
  test_statistic <- t_test$statistic
  p_value <- t_test$p.value
  ci_t <- t_test$conf.int
  
  ## Cohen's d 
  cohend <- cohens_d(lkm_data, pmr_data, ci = 0.95)
  
  cat("\n\n==== Variable:", variable_name, "====\n")
  
  cat("t-test:\n")
  cat("t =", test_statistic, "| p =", p_value, "\n")
  cat("95% CI of mean difference:", ci_t, "\n")
  
  cat("\nCohen's d with 95% CI:\n")
  print(cohend)

}


## Variables
variables <- c("loneliness_mean_T1", "traitEmpathyIRI_EC_mean_T1", "traitEmpathyIRI_PT_mean_T1")

## Run all variables
for (variable in variables) {
  perform_full_analysis(variable, dfB_noT3)
}


```

### 3.2. Correlations at baseline full sample 

```{r, message = FALSE}

# Define variables of interest
selected_vars <- c("socialConnectedness_mean_T1", "loneliness_mean_T1", "inclusionOtherSelf_mean_T1", "traitEmpathyIRI_EC_mean_T1", "traitEmpathyIRI_fant_mean_T1", "traitEmpathyIRI_PD_mean_T1", "wellbeing_mean_T1", "age", "Female", "Male")

# Subset your data frame
subset_data <- dfB_noT3[selected_vars]

# Custom function to calculate p-values for a correlation matrix (Spearman)
cor.mtest <- function(mat, conf.level = 0.95) {
  mat <- as.matrix(mat)
  n <- ncol(mat)
  p.mat <- matrix(NA, n, n)
  diag(p.mat) <- 0
  for (i in 1:(n - 1)) {
    for (j in (i + 1):n) {
      tmp <- cor.test(mat[, i], mat[, j], method = "spearman", exact = FALSE) # Use Spearman
      p.mat[i, j] <- p.mat[j, i] <- tmp$p.value
    }
  }
  colnames(p.mat) <- rownames(p.mat) <- colnames(mat)
  p.mat
}

# Calculate Spearman correlation matrix
cor_matrix <- cor(subset_data, method = "spearman", use = "pairwise.complete.obs")

# Calculate p-values using Spearman correlations
p_matrix <- cor.mtest(subset_data)

# Create a function to format the correlations with asterisks and p-values
format_correlation <- function(cor, p) {
  stars <- ifelse(p < 0.001, "***", ifelse(p < 0.01, "**", ifelse(p < 0.05, "*", "")))
  paste0(sprintf("%.2f", cor), stars, " (", format(p, digits = 2), ")")
}

# Apply the function to format the entire correlation matrix
formatted_cor_matrix <- matrix(NA, nrow = nrow(cor_matrix), ncol = ncol(cor_matrix))
for (i in 1:nrow(cor_matrix)) {
  for (j in 1:ncol(cor_matrix)) {
    if (i != j) {
      formatted_cor_matrix[i, j] <- format_correlation(cor_matrix[i, j], p_matrix[i, j])
    } else {
      formatted_cor_matrix[i, j] <- "-"
    }
  }
}

# Extract the lower triangle of the correlation matrix
lower_triangle <- formatted_cor_matrix
lower_triangle[upper.tri(lower_triangle, diag = TRUE)] <- ""

# Convert lower_triangle to a data frame and add variable names
lower_triangle_df <- as.data.frame(lower_triangle)
colnames(lower_triangle_df) <- selected_vars
rownames(lower_triangle_df) <- selected_vars

# Calculate means and standard deviations
means <- round(sapply(subset_data, mean, na.rm = TRUE), 2)
sds <- round(sapply(subset_data, sd, na.rm = TRUE), 2)

# Create a data frame for descriptive statistics
desc_stats <- data.frame(Variable = selected_vars, Mean = means, SD = sds)

# Combine the descriptive statistics and the lower triangle correlation matrix
desc_stats_with_corr <- cbind(desc_stats, lower_triangle_df)
rownames(desc_stats_with_corr) <- NULL

# Print the final table
print(desc_stats_with_corr)

# save 
write.csv(desc_stats_with_corr, file.path(path, "clean/output/spearman_correlation_table_FullSample_all_with_pvalues_T1.csv"), row.names = FALSE)


# Output the table in a formatted view
kable(desc_stats_with_corr, format = "pandoc", caption = "Descriptive Statistics, Spearman Correlations, and P-values")

```
### 3.2. Correlations at T2 full sample 

```{r, message = FALSE}

# Define variables of interest
selected_vars <- c("socialConnectedness_mean_T2", "loneliness_mean_T2", "inclusionOtherSelf_mean_T2", "traitEmpathyIRI_EC_mean_T2", "traitEmpathyIRI_fant_mean_T2", "traitEmpathyIRI_PD_mean_T2", "traitEmpathyIRI_PT_mean_T2", "wellbeing_mean_T2", "age", "Female", "Male")

# Subset your data frame
subset_data <- dfB_noT3[selected_vars]

# Custom function to calculate p-values for a correlation matrix (Spearman)
cor.mtest <- function(mat, conf.level = 0.95) {
  mat <- as.matrix(mat)
  n <- ncol(mat)
  p.mat <- matrix(NA, n, n)
  diag(p.mat) <- 0
  for (i in 1:(n - 1)) {
    for (j in (i + 1):n) {
      tmp <- cor.test(mat[, i], mat[, j], method = "spearman", exact = FALSE) # Use Spearman
      p.mat[i, j] <- p.mat[j, i] <- tmp$p.value
    }
  }
  colnames(p.mat) <- rownames(p.mat) <- colnames(mat)
  p.mat
}

# Calculate Spearman correlation matrix
cor_matrix <- cor(subset_data, method = "spearman", use = "pairwise.complete.obs")

# Calculate p-values using Spearman correlations
p_matrix <- cor.mtest(subset_data)

# Create a function to format the correlations with asterisks and p-values
format_correlation <- function(cor, p) {
  stars <- ifelse(p < 0.001, "***", ifelse(p < 0.01, "**", ifelse(p < 0.05, "*", "")))
  paste0(sprintf("%.2f", cor), stars, " (", format(p, digits = 2), ")")
}

# Apply the function to format the entire correlation matrix
formatted_cor_matrix <- matrix(NA, nrow = nrow(cor_matrix), ncol = ncol(cor_matrix))
for (i in 1:nrow(cor_matrix)) {
  for (j in 1:ncol(cor_matrix)) {
    if (i != j) {
      formatted_cor_matrix[i, j] <- format_correlation(cor_matrix[i, j], p_matrix[i, j])
    } else {
      formatted_cor_matrix[i, j] <- "-"
    }
  }
}

# Extract the lower triangle of the correlation matrix
lower_triangle <- formatted_cor_matrix
lower_triangle[upper.tri(lower_triangle, diag = TRUE)] <- ""

# Convert lower_triangle to a data frame and add variable names
lower_triangle_df <- as.data.frame(lower_triangle)
colnames(lower_triangle_df) <- selected_vars
rownames(lower_triangle_df) <- selected_vars

# Calculate means and standard deviations
means <- round(sapply(subset_data, mean, na.rm = TRUE), 2)
sds <- round(sapply(subset_data, sd, na.rm = TRUE), 2)

# Create a data frame for descriptive statistics
desc_stats <- data.frame(Variable = selected_vars, Mean = means, SD = sds)

# Combine the descriptive statistics and the lower triangle correlation matrix
desc_stats_with_corr <- cbind(desc_stats, lower_triangle_df)
rownames(desc_stats_with_corr) <- NULL

# Print the final table
print(desc_stats_with_corr)

# save 
write.csv(desc_stats_with_corr, file.path(path, "clean/output/spearman_correlation_table_FullSample_all_with_pvalues_T1.csv"), row.names = FALSE)


# Output the table in a formatted view
kable(desc_stats_with_corr, format = "pandoc", caption = "Descriptive Statistics, Spearman Correlations, and P-values")

```


## 4. Trait Measures: Changes Over Time

### 4.1. Prep time variable so we can look at changes from T1-T2 and T2-T3

```{r, message = FALSE}

# Mean-center GroupID to -0.5 and 0.5
dfB_all_long$GroupID <- factor(dfB_all_long$GroupID, 
                                      levels = c("PMR", "LKM"),  # Original levels
                                      labels = c(-0.5, 0.5))     # New levels

# Check the structure
str(dfB_all_long$GroupID)


# Define the time variable as a factor
dfB_all_long$time <- factor(dfB_all_long$time, levels = c(1, 2, 3))

# Correct the contrast matrix: 3 rows for 3 levels, 2 columns for 2 contrasts
contrasts(dfB_all_long$time) <- t(rbind(
  c(-1, 1, 0),  # Contrast T1 vs T2
  c(0, -1, 1)   # Contrast T2 vs T3
))

# Check the contrasts to ensure they're applied correctly
contrasts(dfB_all_long$time)


# Summary of Advantages of coding GroupID as 0.5 vs -0.5:
# Intercept: Represents the overall mean instead of one group.
# GroupID Coefficient: Symmetric, intuitive, and unaffected by the reference group choice.
# Stability: Reduces multicollinearity, especially with interactions.
# Consistency: Aligns binary predictors with mean-centered continuous variables.
# For our specific case, coding LKM as 0.5 and PMR as -0.5 is particularly beneficial because:
# 
# LKM is experimental group.
# We have interaction terms (e.g., time * GroupID).
# The interpretation becomes cleaner and more symmetrical.

```

### 4.2.1. Loneliness over time

```{r, message = FALSE}

# Run the model with the custom contrasts
fitH1 <- lme(loneliness_mean ~ 1 + time*GroupID + age + Male + Other + YesCollegeDegree, 
                         random = ~1 + time | SubID, 
                         data = dfB_all_long, 
                         method = "ML", 
                         na.action = na.exclude)


tab_model(lme(loneliness_mean ~ 1 + time*GroupID + age + Male + Other + YesCollegeDegree, 
                         random = ~ 1 + time | SubID, 
                         data = dfB_all_long, 
                         method = "ML", 
                         na.action = na.exclude), 
          show.df = T, 
          show.aic = T, 
          show.se = T, 
          show.stat = T, 
          show.std = T)

# save model 
tab_model(fitH1, 
          file = file.path(path, "clean/output/", "fit1H1.html"), 
          show.df = TRUE, 
          show.aic = TRUE, 
          show.se = TRUE, 
          show.stat = TRUE, 
          show.std = TRUE)

# Run the model with the custom contrasts
fitH2 <- lme(socialConnectedness_mean ~ 1 + time*GroupID + age + Male + Other + YesCollegeDegree, 
                         random = ~1 + time | SubID, 
                         data = dfB_all_long, 
                         method = "ML", 
                         na.action = na.exclude)


tab_model(lme(socialConnectedness_mean ~ 1 + time*GroupID + age + Male + Other + YesCollegeDegree, 
                         random = ~ 1 + time | SubID, 
                         data = dfB_all_long, 
                         method = "ML", 
                         na.action = na.exclude), 
          show.df = T, 
          show.aic = T, 
          show.se = T, 
          show.stat = T, 
          show.std = T)

# save model 
tab_model(fitH2, 
          file = file.path(path, "clean/output/", "fit1H2.html"), 
          show.df = TRUE, 
          show.aic = TRUE, 
          show.se = TRUE, 
          show.stat = TRUE, 
          show.std = TRUE)

```

### 4.2.2. Bayes Factors: Loneliness and Social Connectedness


```{r, message = FALSE}

# Rouder, J. N., Speckman, P. L., Sun, D., Morey, R. D., & Iverson, G. (2009). Bayesian t tests for accepting and rejecting the null hypothesis. Psychonomic Bulletin & Review, 16(2), 225–237. https://doi.org/10.3758/PBR.16.2.225


# LONELINESS
# Data
t_values <- c(-1.01, 1.86, 1.16, 0.62)
dfs <- c(201, 102, 201, 201)
labels <- c("time2", "Group", "time1:Group", "time2:Group")

# Compute BF10 correctly (log BF → exponentiated)
BFs <- mapply(function(t, df) {
  res <- BayesFactor::ttest.tstat(t = t, n1 = df + 1, rscale = "medium")
  bf <- exp(res$bf)  # fix: exponentiate the log(BF10)
  return(bf)
}, t_values, dfs)

# Interpretation labels
BF_labels <- sapply(BFs, function(bf) {
  if (bf < 1/10) return("Strong evidence for H0")
  else if (bf < 1/3) return("Moderate evidence for H0")
  else if (bf < 1) return("Anecdotal evidence for H0")
  else if (bf < 3) return("Anecdotal evidence for H1")
  else if (bf < 10) return("Moderate evidence for H1")
  else return("Strong evidence for H1")
})

# Results table
results <- data.frame(
  Effect = labels,
  t_value = t_values,
  df = dfs,
  BF10 = round(BFs, 3),
  BF01 = round(1 / BFs, 3),
  Interpretation = BF_labels
)

print(results)


# SOCIAL CONNECTEDNESS 

library(BayesFactor)

# Data
t_values <- c(1.85, 1.18, -1.75, -1.12, -1.40)
dfs <- c(201, 201, 102, 201, 201)
labels <- c("time1","time2", "Group", "time1:Group", "time2:Group")

# Compute BF10 correctly (log BF → exponentiated)
BFs <- mapply(function(t, df) {
  res <- BayesFactor::ttest.tstat(t = t, n1 = df + 1, rscale = "medium")
  bf <- exp(res$bf)  # fix: exponentiate the log(BF10)
  return(bf)
}, t_values, dfs)

# Interpretation labels
BF_labels <- sapply(BFs, function(bf) {
  if (bf < 1/10) return("Strong evidence for H0")
  else if (bf < 1/3) return("Moderate evidence for H0")
  else if (bf < 1) return("Anecdotal evidence for H0")
  else if (bf < 3) return("Anecdotal evidence for H1")
  else if (bf < 10) return("Moderate evidence for H1")
  else return("Strong evidence for H1")
})

# Results table
results <- data.frame(
  Effect = labels,
  t_value = t_values,
  df = dfs,
  BF10 = round(BFs, 3),
  BF01 = round(1 / BFs, 3),
  Interpretation = BF_labels
)

print(results)


# yields similar results: 
# Wagenmakers, E.-J. (2007). A practical solution to the pervasive problems of p values. Psychonomic Bulletin & Review, 14(5), 779–804. https://doi.org/10.3758/BF03194105

# # Function to calculate BF10 from t and df according to EJs paper
# BF10_from_t <- function(t, df) {
#   n <- df + 1
#   BF01 <- sqrt(n) * exp(- (t^2) / 2)
#   BF10 <- 1 / BF01
#   return(BF10)
# }
# 
# # Your t-values & dfs
# t_values <- c(-1.01, 1.86, 1.16, 0.62)
# dfs <- c(201, 102, 201, 201)
# 
# # Labels
# effects <- c("time2", "Group", "time1:Group", "time2:Group")
# 
# # Calculate BFs
# BFs <- BF10_from_t(t_values, dfs)
# 
# # Show nicely
# data.frame(
#   Effect = effects,
#   t_value = t_values,
#   df = dfs,
#   BF10 = round(BFs, 3)
# )


```


### 4.2.5. Plot loneliness

```{r, message = FALSE}
library(ggplot2)
library(dplyr)
library(ggExtra)
library(scico)

# ===== knobs =====
gap <- 0.10        # smaller = closer T1–T3
density_size <- 8  # lower = thinner density panel (try 3–5)
density_alpha <- 0.6
aspect_ratio <- 0.7  # >1 = taller & thinner overall look
# ==================

# Prepare data
df_tmp <- dfB_all_long %>%
  mutate(
    GroupID_label = ifelse(GroupID == 0.5, "LKM", "PMR"),
    time_chr = case_when(
      is.numeric(time)        ~ paste0("T", as.integer(time)),
      grepl("^T?\\d+$", time) ~ ifelse(grepl("^T", time), time, paste0("T", time)),
      TRUE                    ~ as.character(time)
    )
  )

levs <- c("T1","T2","T3")
pos  <- setNames(1 + (0:(length(levs)-1))*gap, levs)

loneliness_data <- df_tmp %>%
  mutate(
    time_f   = factor(time_chr, levels = levs),
    time_num = unname(pos[as.character(time_chr)])
  ) %>%
  filter(!is.na(time_num), !is.na(loneliness_mean)) %>%
  select(SubID, time_f, time_num, GroupID_label, loneliness_mean)

# Shared y range
ymin <- min(loneliness_data$loneliness_mean, na.rm = TRUE)
ymax <- max(loneliness_data$loneliness_mean, na.rm = TRUE)
yrange <- c(ymin - 0.05, ymax + 0.05)

# Main plot
p_main <- ggplot(
  loneliness_data,
  aes(x = time_num, y = loneliness_mean,
      group = GroupID_label, color = GroupID_label)
) +
  stat_summary(fun = mean, geom = "line", aes(linetype = GroupID_label), size = 1.1) +
  stat_summary(fun.data = mean_se, geom = "errorbar",
               width = gap/6, size = 0.4, color = "black") +  # <<< smaller SE bars
  geom_jitter(width = gap/4, alpha = 0.4, size = 1.4, show.legend = FALSE) +
  papaja::theme_apa() +
  labs(
    x = "Time",
    y = "Loneliness Mean Score",
    linetype = "Group",
    color = "Group"
  ) +
  scale_color_manual(values = c("hotpink", "blue")) + scale_x_continuous(
    breaks = unname(pos), labels = names(pos),
    limits = c(min(pos) - gap*0.5, max(pos) + gap*0.5),
    expand = expansion(mult = c(0, 0))
  ) +
  coord_cartesian(ylim = yrange) +
  theme(
    legend.position = "bottom",
    plot.margin = margin(2, 2, 2, 2),
    aspect.ratio = aspect_ratio
  )

# Add slim right-side density aligned with y
p_final <- ggMarginal(
  p_main,
  type = "density",
  margins = "y",
  size = density_size,
  groupColour = TRUE,
  groupFill = TRUE,
  alpha = density_alpha
)

print(p_final)




```

### 4.3.1. IRI and subscales over time

```{r, message = FALSE}

# Run the model with the custom contrasts
fitH1e_1_contrast <- lme(traitEmpathyIRI_EC_mean ~ 1 + time*GroupID + age + Male + Other + YesCollegeDegree, 
                         random = ~1 + time | SubID, 
                         data = dfB_all_long, 
                         method = "ML", 
                         na.action = na.exclude)


tab_model(lme(traitEmpathyIRI_EC_mean ~ 1 + time*GroupID + age + Male + Other + YesCollegeDegree, 
                         random = ~1 + time | SubID, 
                         data = dfB_all_long, 
                         method = "ML", 
                         na.action = na.exclude), 
          show.df = T, 
          show.aic = T, 
          show.se = T, 
          show.stat = T, 
          show.std = T)

# save model 
tab_model(fitH1e_1_contrast, file = "fit1H1e_T3.html", 
          show.df = T, 
          show.aic = T, 
          show.se = T, 
          show.stat = T, 
          show.std = T)

# Run the model with the custom contrasts
fitH1g_1_contrast <- lme(traitEmpathyIRI_PT_mean ~ 1 + time*GroupID + age + Male + Other + YesCollegeDegree, 
                         random = ~1 + time | SubID, 
                         data = dfB_all_long, 
                         method = "ML", 
                         na.action = na.exclude)


tab_model(lme(traitEmpathyIRI_PT_mean ~ 1 + time*GroupID + age + Male + Other + YesCollegeDegree, 
                         random = ~1 + time | SubID, 
                         data = dfB_all_long, 
                         method = "ML", 
                         na.action = na.exclude), 
          show.df = T, 
          show.aic = T, 
          show.se = T, 
          show.stat = T, 
          show.std = T)

# save model 
tab_model(fitH1g_1_contrast, file = "fit1H1g_T3.html", 
          show.df = T, 
          show.aic = T, 
          show.se = T, 
          show.stat = T, 
          show.std = T)


```

### 4.3.2. Bayes Factors: IRI and subscales over time

```{r, message = FALSE}

library(BayesFactor)

# === 1. t-values and dfs for IRI EC model ===
t_values_EC <- c(-0.12, -1.00, -0.29, 1.10, 1.18)
dfs_EC <- c(201, 201, 102, 201, 201)
effects_EC <- c("time1", "time2", "Group",
                "time1:Group", "time2:Group")

# === 2. t-values and dfs for IRI PT model ===
t_values_PT <- c(1.04, 0.13, -0.02, 0.64, 0.68)
dfs_PT <- c(201, 201, 102, 201, 201)
effects_PT <- c("time1", "time2", "Group", 
                "time1:Group", "time2:Group")

# === 3. Function to calculate BF10 from t and df ===
compute_BF_table <- function(t_vals, dfs, labels) {
  BFs <- mapply(function(t, df) {
    res <- BayesFactor::ttest.tstat(t = t, n1 = df + 1, rscale = "medium")
    exp(res$bf)  # exponentiate the log(BF10)
  }, t_vals, dfs)
  
  BF_labels <- sapply(BFs, function(bf) {
    if (bf < 1/10) return("Strong evidence for H0")
    else if (bf < 1/3) return("Moderate evidence for H0")
    else if (bf < 1) return("Anecdotal evidence for H0")
    else if (bf < 3) return("Anecdotal evidence for H1")
    else if (bf < 10) return("Moderate evidence for H1")
    else return("Strong evidence for H1")
  })
  
  data.frame(
    Effect = labels,
    t_value = t_vals,
    df = dfs,
    BF10 = round(BFs, 3),
    BF01 = round(1 / BFs, 3),
    Interpretation = BF_labels
  )
}

# === 4. Compute tables ===
results_EC <- compute_BF_table(t_values_EC, dfs_EC, effects_EC)
results_PT <- compute_BF_table(t_values_PT, dfs_PT, effects_PT)

# === 5. Print results ===
cat("===== IRI EC Model =====\n")
print(results_EC)

cat("\n===== IRI PT Model =====\n")
print(results_PT)

```

### 4.4.1. IOS mean and stranger over time 

```{r, message = FALSE}

# Run the model with the custom contrasts

tab_model(lme(inclusionOtherSelf_mean ~ 1 + time*GroupID + age + Male + Other + YesCollegeDegree, 
                         random = ~1 + time | SubID, 
                         data = dfB_all_long, 
                         method = "ML", 
                         na.action = na.exclude), 
          show.df = T, 
          show.aic = T, 
          show.se = T, 
          show.stat = T, 
          show.std = T)


tab_model(lme(inclusionOtherSelf_mean ~ 1 + time*GroupID + age + Male + Other + YesCollegeDegree, 
                         random = ~1 + time | SubID, 
                         data = dfB_all_long, 
                         method = "ML", 
                         na.action = na.exclude), 
          show.df = T, 
          show.aic = T, 
          show.se = T, 
          show.stat = T, 
          show.std = T)

# Run the model with the custom contrasts
fitH1h_1_contrast <- lme(IOS_strangerNum ~ 1 + time*GroupID + age + Male + Other + YesCollegeDegree, 
                         random = ~1 + time | SubID, 
                         data = dfB_all_long, 
                         method = "ML", 
                         na.action = na.exclude)


tab_model(lme(IOS_strangerNum ~ 1 + time*GroupID + age + Male + Other + YesCollegeDegree, 
                         random = ~1 + time | SubID, 
                         data = dfB_all_long, 
                         method = "ML", 
                         na.action = na.exclude), 
          show.df = T, 
          show.se = T, 
          show.stat = T, 
          show.std = T)


fit_ios_stranger <- lme(IOS_strangerNum ~ 1 + time*GroupID + age + Male + Other + YesCollegeDegree, 
                         random = ~1 + time | SubID, 
                         data = dfB_all_long, 
                         method = "ML", 
                         na.action = na.exclude)

# save model 
tab_model(fit_ios_stranger, file = "fit_ios_stranger.html", 
          show.df = T, 
          show.se = T, 
          show.stat = T, 
          show.std = T)


```

### 4.4.2. Bayes Factors: IOS mean and stranger over time 

```{r, message = FALSE}

library(BayesFactor)

# === 1. t-values and dfs for IOS mean ===
t_values_ios_mean <- c(0.66, -0.59, 0.47, 1.44)
dfs_ios_mean <- c(201, 102, 201, 201)
effects_ios_mean <- c("time2", "Group", "time1:Group", "time2:Group")

# === 2. t-values and dfs for IOS stranger (no demographics) ===
t_values_ios_stranger <- c(0.67, 0.09, -0.95, -0.39)
dfs_ios_stranger <- c(201, 102, 201, 201)
effects_ios_stranger <- c("time2", "Group", "time1:Group", "time2:Group")

# === 3. Function to compute BF10 with interpretation ===
compute_BF_table <- function(t_vals, dfs, labels) {
  BFs <- mapply(function(t, df) {
    res <- BayesFactor::ttest.tstat(t = t, n1 = df + 1, rscale = "medium")
    exp(res$bf)
  }, t_vals, dfs)
  
  BF_labels <- sapply(BFs, function(bf) {
    if (bf < 1/10) return("Strong evidence for H0")
    else if (bf < 1/3) return("Moderate evidence for H0")
    else if (bf < 1) return("Anecdotal evidence for H0")
    else if (bf < 3) return("Anecdotal evidence for H1")
    else if (bf < 10) return("Moderate evidence for H1")
    else return("Strong evidence for H1")
  })
  
  data.frame(
    Effect = labels,
    t_value = t_vals,
    df = dfs,
    BF10 = round(BFs, 3),
    BF01 = round(1 / BFs, 3),
    Interpretation = BF_labels
  )
}

# === 4. Compute tables ===
results_ios_mean <- compute_BF_table(t_values_ios_mean, dfs_ios_mean, effects_ios_mean)
results_ios_stranger <- compute_BF_table(t_values_ios_stranger, dfs_ios_stranger, effects_ios_stranger)

# === 5. Print results ===
cat("===== Inclusion Other Self Mean (IOS mean) — No Demographics =====\n")
print(results_ios_mean)

cat("\n===== IOS Stranger Num — No Demographics =====\n")
print(results_ios_stranger)


```

### 4.4. Plot IOS mean and stranger

```{r, message = FALSE}
# Load required packages
library(ggplot2)
library(tidyr)
library(dplyr)
library(papaja)

# Reshape the data to long format for all 4 subscales
df_long_ios <- dfB_all_long %>%
  select(SubID, time, GroupID, inclusionOtherSelf_mean, IOS_strangerNum) %>%
  pivot_longer(cols = c(inclusionOtherSelf_mean, IOS_strangerNum),
               names_to = "measure",
               values_to = "value")

# Define labels for the measures
df_long_ios$measure <- factor(df_long_ios$measure, 
                          levels = c(
                                     "inclusionOtherSelf_mean", "IOS_strangerNum"),
                          labels = c("IOS Mean", "IOS Stranger"))

# Update GroupID labels
df_long_ios <- df_long_ios %>%
  mutate(GroupID = recode(GroupID, `-0.5` = "PMR", `0.5` = "LKM"))

# Create the main plot with time converted to a factor
p <- ggplot(df_long_ios, aes(x = factor(time), y = value, group = interaction(measure, GroupID))) +
  stat_summary(fun = mean, geom = "line", aes(color = measure, linetype = GroupID), size = 1.2) +  # Different line types for GroupIDs, colors for measure
  stat_summary(fun = mean, geom = "point", size = 3, aes(shape = GroupID), position = position_dodge(width = 0.2)) +  # Add points for each GroupID
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.15, size = 0.8, 
               position = position_dodge(width = 0.2)) +  # Add error bars with dodge
  geom_jitter(width = 0.1, alpha = 0.3, aes(color = measure, shape = GroupID), size = 1.5, show.legend = FALSE) +  # Add scatter points for each GroupID
  theme_apa() +  # Use a clean, minimal theme
  labs(x = "Time",
       y = "Mean Score",
       color = "Subscale",
       linetype = "Meditation Type",
       shape = "Meditation Type") + 
  scale_x_discrete(labels = c("T1", "T2", "T3")) +  # Label time points
  scale_color_manual(values = c("salmon", "violet")) +  # Use custom colors for subscales
    scale_linetype_manual(values = c("PMR" = "twodash", "LKM" = "solid")) +  # Specify linetypes for GroupID

  guides(color = guide_legend(override.aes = list(linetype = 1, size = 1.2)),  # Override aesthetics for measure to show only lines
         linetype = guide_legend(order = 1),  # Ensure linetype is in the right order
         shape = guide_legend(order = 1)) +  # Ensure shape is in the right order
  theme(legend.position = "right",
        plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 10),
        legend.key.width = unit(0.9, "cm"))

# Display the plot
print(p)


```

## 5. Differences in subjetive ratings

### 5.1. 2sample T-tests

```{r, message = FALSE}

#detach("package:plyr", unload = TRUE)

# Load necessary libraries
library(dplyr)
library(ggplot2)
library(tidyr)  # needed for pivot_longer

# Define your results data frame (t-test results remain unchanged)
results <- data.frame(Variable = character(), 
                      Mean_LKM = numeric(), 
                      Mean_PMR = numeric(), 
                      SD_LKM = numeric(), 
                      SD_PMR = numeric(), 
                      t_value = numeric(), 
                      df = numeric(), 
                      p_value = numeric(), 
                      CI_lower = numeric(),
                      CI_upper = numeric(),
                      Significance = character(),
                      Effect_Size = numeric(),
                      stringsAsFactors = FALSE)

# Variables and their corresponding labels (using your original list)
variables <- c("PSI", "painRatingSelf", "self_fearOfPain", "self_unpleasantPain", 
               "painRatingOther", "other_fearOfPain", "other_unpleasantPain")

labels <- c("PSI", 
            "Pain Rating (Self)", 
            "Fear Rating (Self)", 
            "Unpleasantness Rating (Self)", 
            "Pain Rating (Other)", 
            "Fear Rating (Other)", 
            "Unpleasantness Rating (Other)")

# Loop through each variable to perform t-tests on raw data
for (var in variables) {
  # Perform independent 2-sample t-test
  t_test <- t.test(dfS_noT3[[var]] ~ dfS_noT3$GroupID, var.equal = TRUE)
  
  # Calculate means and SDs for each group
  means <- dfS_noT3 %>% 
    group_by(GroupID) %>% 
    summarise(
      Mean = mean(get(var), na.rm = TRUE),
      SD = sd(get(var), na.rm = TRUE)
    )
  
  # Calculate effect size (Cohen's d)
  n_lkm <- sum(dfS_noT3$GroupID == "LKM")
  n_pmr <- sum(dfS_noT3$GroupID == "PMR")
  pooled_sd <- sqrt(((n_lkm - 1) * means$SD[means$GroupID == "LKM"]^2 + 
                     (n_pmr - 1) * means$SD[means$GroupID == "PMR"]^2) / 
                    (n_lkm + n_pmr - 2))
  cohen_d <- abs(means$Mean[means$GroupID == "LKM"] - means$Mean[means$GroupID == "PMR"]) / pooled_sd
  
  # Append results to the data frame with CI bounds
  results <- rbind(results, data.frame(
    Variable = var,
    Mean_LKM = means$Mean[means$GroupID == "LKM"],
    Mean_PMR = means$Mean[means$GroupID == "PMR"],
    SD_LKM = means$SD[means$GroupID == "LKM"],
    SD_PMR = means$SD[means$GroupID == "PMR"],
    t_value = t_test$statistic,
    df = t_test$parameter,
    p_value = t_test$p.value,
    CI_lower = t_test$conf.int[1],
    CI_upper = t_test$conf.int[2],
    Significance = ifelse(t_test$p.value < 0.05, "Significant", "Not Significant"),
    Effect_Size = cohen_d
  ))
}

# Save the results table as a CSV file (APA appropriate)
results_apa <- results %>% mutate_if(is.numeric, ~ round(., 2))
write.csv(results_apa, file.path(path, "clean/output/APA_2sample_t_test_subjectiveRatings_results.csv"), row.names = FALSE)

print(results_apa)

# For plotting, standardize (z-score) all the variables
# Create new standardized columns named "std_<variable>"
dfS_noT3_std <- dfS_noT3 %>%
  mutate(across(all_of(variables), ~ as.numeric(scale(.)), .names = "std_{col}"))

# Create a vector of standardized variable names
std_variables <- paste0("std_", variables)

# Create a long-format data frame for plotting using the standardized variables
df_long_std <- dfS_noT3_std %>%
  select(GroupID, all_of(std_variables)) %>%
  pivot_longer(cols = -GroupID, names_to = "Variable", values_to = "Value") %>%
  mutate(Variable = factor(Variable, levels = std_variables, labels = labels))

# Define custom colors: adjust as needed (example: purple and blue)
custom_colors <- c("LKM" = "#377EB8",  # Purple
                   "PMR" = "#4DA")  # Orange
# Create one combined violin plot using the standardized values
plot_all <- ggplot(df_long_std, aes(x = Variable, y = Value, fill = GroupID)) +
  geom_violin(position = position_dodge(width = 0.8), alpha = 0.6) +
  geom_jitter(aes(fill = GroupID), 
              shape = 21,                # circle with fill and outline
              color = "black",            # grey outline
              position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.8),
              size = 1.5, alpha = 0.4) +
  stat_summary(fun = mean, geom = "point", position = position_dodge(width = 0.8), 
               shape = 21, size = 3, fill = "white") +
  stat_summary(fun.data = mean_se, geom = "errorbar", position = position_dodge(width = 0.8), 
               width = 0.2) +
  labs(title = "Comparison of Standardized Variables",
       x = "Variable", y = "Standardized Value (z-score)") +
  scale_fill_manual(values = custom_colors) +
  scale_color_manual(values = custom_colors) +
  theme_apa() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Display the combined plot
print(plot_all)



```
### 5.2. Study partner 

```{r, message = FALSE}
# Load necessary libraries
library(dplyr)
library(ggplot2)
library(tidyr)   # for pivot_longer

# Define your results data frame
results <- data.frame(Variable = character(), 
                      Mean_LKM = numeric(), 
                      Mean_PMR = numeric(), 
                      SD_LKM = numeric(), 
                      SD_PMR = numeric(), 
                      t_value = numeric(), 
                      df = numeric(), 
                      p_value = numeric(), 
                      CI_lower = numeric(),
                      CI_upper = numeric(),
                      Significance = character(),
                      Effect_Size = numeric(),
                      stringsAsFactors = FALSE)

# Define variables and their labels in the desired order:
# 1. IOS (Mean) from inclusionOtherSelf_mean_T2
# 2. IOS (Stranger) from IOS_strangerNum_T2
# 3. IOS (Study Partner) from IOS_studyPartner
# 4. Liking (Study Partner) from studyPartnerLiking
# 5. Real (Study Partner) from studyPartnerReal
variables <- c("inclusionOtherSelf_mean_T2", "IOS_strangerNum_T2", 
               "IOS_studyPartner", "studyPartnerLiking", "studyPartnerReal")
labels <- c("IOS (Mean)", "IOS (Stranger)", "IOS (Study Partner)", 
            "Liking (Study Partner)", "Real (Study Partner)")

# Loop through each variable to perform t-tests on raw data
for (var in variables) {
  # Perform independent 2-sample t-test
  t_test <- t.test(dfS_noT3[[var]] ~ dfS_noT3$GroupID, var.equal = TRUE)
  
  # Calculate means and SDs for each group
  means <- dfS_noT3 %>% 
    group_by(GroupID) %>% 
    summarise(
      Mean = mean(.data[[var]], na.rm = TRUE),
      SD = sd(.data[[var]], na.rm = TRUE)
    )
  
  # Calculate effect size (Cohen's d)
  n_lkm <- sum(dfS_noT3$GroupID == "LKM")
  n_pmr <- sum(dfS_noT3$GroupID == "PMR")
  pooled_sd <- sqrt(((n_lkm - 1) * means$SD[means$GroupID == "LKM"]^2 + 
                     (n_pmr - 1) * means$SD[means$GroupID == "PMR"]^2) / 
                    (n_lkm + n_pmr - 2))
  cohen_d <- abs(means$Mean[means$GroupID == "LKM"] - means$Mean[means$GroupID == "PMR"]) / pooled_sd
  
  # Append results to the data frame with CI bounds
  results <- rbind(results, data.frame(
    Variable = var,
    Mean_LKM = means$Mean[means$GroupID == "LKM"],
    Mean_PMR = means$Mean[means$GroupID == "PMR"],
    SD_LKM = means$SD[means$GroupID == "LKM"],
    SD_PMR = means$SD[means$GroupID == "PMR"],
    t_value = t_test$statistic,
    df = t_test$parameter,
    p_value = t_test$p.value,
    CI_lower = t_test$conf.int[1],
    CI_upper = t_test$conf.int[2],
    Significance = ifelse(t_test$p.value < 0.05, "Significant", "Not Significant"),
    Effect_Size = cohen_d
  ))
}

# Round numeric columns for APA-style presentation and save results as CSV
results_apa <- results %>% mutate_if(is.numeric, ~ round(., 2))
write.csv(results_apa, file.path(path, "clean/output/APA_2sample_t_test_studyPartner_results.csv"), row.names = FALSE)
print(results_apa)

# For plotting, standardize (z-score) all the variables for better comparability.
# Create new standardized columns (named std_<variable>)
dfS_noT3_std <- dfS_noT3 %>%
  mutate(across(all_of(variables), ~ as.numeric(scale(.)), .names = "std_{col}"))

# Create a vector of standardized variable names in the same order
std_variables <- paste0("std_", variables)

# Create a long-format data frame for plotting using the standardized variables
df_long_std <- dfS_noT3_std %>%
  select(GroupID, all_of(std_variables)) %>%
  pivot_longer(cols = -GroupID, names_to = "Variable", values_to = "Value") %>%
  mutate(Variable = factor(Variable, levels = std_variables, labels = labels))

# Define custom colors: Purple for "LKM" and Orange for "PMR"
custom_colors <- c("LKM" = "#800080",  # Purple
                   "PMR" = "pink")  # Orange

# Create the violin plot with jitter (jitter points using shape 21 with grey outlines)
plot_std <- ggplot(df_long_std, aes(x = Variable, y = Value, fill = GroupID)) +
  geom_violin(position = position_dodge(width = 0.8), alpha = 0.6) +
  geom_jitter(aes(fill = GroupID), 
              shape = 21,                 # circle with both fill and outline
              color = "black",             # grey outline
              position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.8),
              size = 1.5, alpha = 0.4) +   # transparent jitter points
  stat_summary(fun = mean, geom = "point", position = position_dodge(width = 0.8),
               shape = 21, size = 3, fill = "white") +
  stat_summary(fun.data = mean_se, geom = "errorbar", position = position_dodge(width = 0.8),
               width = 0.2) +
  labs(title = "Comparison of Standardized Study Partner Variables",
       x = "Variable", y = "Standardized Value (z-score)") +
  scale_fill_manual(values = custom_colors) +
  scale_color_manual(values = custom_colors) +
  theme_apa() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Display the plot
print(plot_std)

```

## 6. State Empathy 

### 6.1. Trait Empathy 

```{r, message = FALSE}

# fear 
tab_model(lm(traitEmpathyIRI_EC_mean_T2 ~ age + Male + Other + YesCollegeDegree + GroupID + stateEmp_fear, data = dfS_noT3), 
          show.df = T, 
          show.aic = T, 
          show.se = T, 
          show.stat = T, 
          show.std = T)


tab_model(lm(traitEmpathyIRI_PT_mean_T2 ~ age + Male + Other + YesCollegeDegree + GroupID + stateEmp_fear, data = dfS_noT3), 
          show.df = T, 
          show.aic = T, 
          show.se = T, 
          show.stat = T, 
          show.std = T)

# pain rating 
tab_model(lm(traitEmpathyIRI_EC_mean_T2 ~ age + Male + Other + YesCollegeDegree + GroupID + stateEmp_painRating, data = dfS_noT3), 
          show.df = T, 
          show.aic = T, 
          show.se = T, 
          show.stat = T, 
          show.std = T)

tab_model(lm(traitEmpathyIRI_PT_mean_T2 ~ age + Male + Other + YesCollegeDegree + GroupID + stateEmp_painRating, data = dfS_noT3), 
          show.df = T, 
          show.aic = T, 
          show.se = T, 
          show.stat = T, 
          show.std = T)

# unpleasantness 
tab_model(lm(traitEmpathyIRI_EC_mean_T2 ~ age + Male + Other + YesCollegeDegree + GroupID + stateEmp_painUnpleasant, data = dfS_noT3), 
          show.df = T, 
          show.aic = T, 
          show.se = T, 
          show.stat = T, 
          show.std = T)

tab_model(lm(traitEmpathyIRI_PT_mean_T2 ~ age + Male + Other + YesCollegeDegree + GroupID + stateEmp_painUnpleasant, data = dfS_noT3), 
          show.df = T, 
          show.aic = T, 
          show.se = T, 
          show.stat = T, 
          show.std = T)




```

### 6.2. Loneliness 

```{r, message = FALSE}

# fear

tab_model(lm(loneliness_mean_T2 ~ age + Male + Other + YesCollegeDegree + GroupID + stateEmp_fear, data = dfS_noT3), 
          show.df = T, 
          show.aic = T, 
          show.se = T, 
          show.stat = T, 
          show.std = T)

tab_model(lm(socialConnectedness_mean_T2 ~ age + Male + Other + YesCollegeDegree + GroupID + stateEmp_fear, data = dfS_noT3), 
          show.df = T, 
          show.aic = T, 
          show.se = T, 
          show.stat = T, 
          show.std = T)

# pain rating 

tab_model(lm(loneliness_mean_T2 ~ age + Male + Other + YesCollegeDegree + GroupID + stateEmp_painRating, data = dfS_noT3), 
          show.df = T, 
          show.aic = T, 
          show.se = T, 
          show.stat = T, 
          show.std = T)

tab_model(lm(socialConnectedness_mean_T2 ~ age + Male + Other + YesCollegeDegree + GroupID + stateEmp_painRating, data = dfS_noT3), 
          show.df = T, 
          show.aic = T, 
          show.se = T, 
          show.stat = T, 
          show.std = T)

# unpleasantness 

tab_model(lm(loneliness_mean_T2 ~ age + Male + Other + YesCollegeDegree + GroupID + stateEmp_painUnpleasant, data = dfS_noT3), 
          show.df = T, 
          show.aic = T, 
          show.se = T, 
          show.stat = T, 
          show.std = T)

tab_model(lm(socialConnectedness_mean_T2 ~ age + Male + Other + YesCollegeDegree + GroupID + stateEmp_painUnpleasant, data = dfS_noT3), 
          show.df = T, 
          show.aic = T, 
          show.se = T, 
          show.stat = T, 
          show.std = T)


# unpleasantness 

tab_model(lm(loneliness_mean_T2 ~ age + Male + Other + YesCollegeDegree + GroupID + traitEmpathyIRI_EC_mean_T2, data = dfS_noT3), 
          show.df = T, 
          show.aic = T, 
          show.se = T, 
          show.stat = T, 
          show.std = T)

tab_model(lm(socialConnectedness_mean_T2 ~  traitEmpathyIRI_PT_mean_T2, data = dfS_noT3), 
          show.df = T, 
          show.aic = T, 
          show.se = T, 
          show.stat = T, 
          show.std = T)

```

#### 6.2. Plots

regression plot

```{r, message = FALSE}

# ============================
# Scatterplot matrix (tight) with:
#  - Upper: standardized β (with covariates)
#  - Lower: red sig stars on the regression line
#  - Diagonal: histograms with labels
# ============================

library(GGally)
library(ggplot2)
library(dplyr)

# --------------------------
# Variables, labels, covariates
# --------------------------
vars <- c(
  "loneliness_mean_T2",
  "socialConnectedness_mean_T2",
  "stateEmp_painRating",
  "stateEmp_painUnpleasant",
  "stateEmp_fear",
  "painRatingOther",
  "other_fearOfPain",
  "other_unpleasantPain"
  
)

labels_pretty <- c(
  loneliness_mean_T2          = "Loneliness",
  socialConnectedness_mean_T2 = "Social Connectedness",
  stateEmp_painRating         = "State Empathy (Pain)",
  stateEmp_painUnpleasant     = "State Empathy (Unpleasant)",
  stateEmp_fear               = "State Empathy (Fear)",
  painRatingSelf              = "Self Rating (Pain)",
  painRatingOther             = "Other Rating (Pain)",
  self_unpleasantPain         = "Self Rating (Unpleasant)",
  other_unpleasantPain        = "Other Rating (Unpleasant)",
  self_fearOfPain             = "Self Rating (Fear)",
  other_fearOfPain            = "Other Rating (Fear)"
)

# Categorical/binary covariates (leave as-is)
covars_cat  <- c("Male","Other","YesCollegeDegree","GroupID")
# Continuous covariates to mean-center
covars_cont <- c("age")

# Build analysis frame and mean-center continuous covariates
dat <- dfS_noT3 |>
  dplyr::select(all_of(unique(c(vars, covars_cat, covars_cont)))) |>
  mutate(across(all_of(covars_cont),
                ~ as.numeric(scale(., center = TRUE, scale = FALSE)),
                .names = "{.col}_c")) |>
  drop_na()

# Use centered names in models
covars <- c(paste0(covars_cont, "_c"), covars_cat)

# --------------------------
# Compact aesthetics
# --------------------------
SIZE_PT   <- 0.9 # points
SIZE_LINE <- 0.4   # lm line
SIZE_TXT  <- 2.3   # upper β text
SIZE_DIAG <- 7   # diagonal labels
SIZE_STAR <- 3.5   # lower-star size (a bit bigger, as requested)
BINS      <- 7

starify <- function(p) ifelse(p < .001,"***",
                         ifelse(p < .01,"**",
                         ifelse(p < .05,"*","")))

# --------------------------
# Panel functions
# --------------------------

# Upper: standardized beta for X in scale(Y) ~ scale(X) + covars
upper_beta <- function(data, mapping, ...){
  yv <- as_label(mapping$y)
  xv <- as_label(mapping$x)

  rhs <- paste0("scale(", xv, ")",
                if (length(covars)) paste0(" + ", paste(covars, collapse = " + ")) else "")
  fml <- as.formula(paste0("scale(", yv, ") ~ ", rhs))

  fit <- try(lm(fml, data = data, na.action = na.exclude), silent = TRUE)

  lab <- "β = NA"
  if (!inherits(fit, "try-error")) {
    nm <- paste0("scale(", xv, ")")
    sm <- summary(fit)$coefficients
    if (nm %in% rownames(sm)) {
      b    <- unname(sm[nm, "Estimate"])
      pval <- unname(sm[nm, "Pr(>|t|)"])
      lab  <- sprintf("\u03b2 = %.2f %s", b, starify(pval))
    }
  }

  ggplot() +
    annotate("text", x = 0.5, y = 0.5, label = lab,
             size = SIZE_TXT, fontface = "bold") +
    theme_void()
}

# Lower: scatter + LM line; red sig stars ON the fitted line at its midpoint
lower_scatter_with_stars <- function(data, mapping, ...){
  x <- GGally::eval_data_col(data, mapping$x)
  y <- GGally::eval_data_col(data, mapping$y)

  # Midpoint of the panel (in data coords)
  xr <- range(x, na.rm = TRUE)
  yr <- range(y, na.rm = TRUE)
  xmid <- mean(xr)

  fit <- try(lm(y ~ x), silent = TRUE)
  stars <- ""
  y_on_line <- mean(yr)  # fallback

  if (!inherits(fit, "try-error")) {
    sm <- summary(fit)$coefficients
    if ("x" %in% rownames(sm)) {
      pval     <- unname(sm["x", "Pr(>|t|)"])
      stars    <- ifelse(pval < .001,"***",
                   ifelse(pval < .01,"**",
                   ifelse(pval < .05,"*","")))
      y_on_line <- as.numeric(predict(fit, newdata = data.frame(x = xmid)))
    }
  }

  ggplot(data, mapping) +
    geom_point(size = SIZE_PT, alpha = 0.75, color = "#E86A33") +
    geom_smooth(method = "lm", se = FALSE, linewidth = SIZE_LINE, color = "grey30") +
    # place stars at the CENTER on the fitted line
    annotate("text", x = xmid, y = y_on_line, label = stars,
             vjust = -0.2, size = SIZE_STAR, fontface = "bold", color = "blue") +
    # small padding so no points/line get clipped by the panel box
    scale_x_continuous(expand = expansion(mult = c(0.06, 0.06))) +
    scale_y_continuous(expand = expansion(mult = c(0.06, 0.06))) +
    coord_cartesian(clip = "on") +
    theme_minimal(base_size = 7) +
    theme(
      panel.grid = element_blank(),
      axis.title = element_blank(),
      axis.text  = element_text(size = 5),
      axis.ticks = element_line(linewidth = 0.2)
    )
}

# Diagonal: compact histogram + label inside
diag_hist_label <- function(data, mapping, ...){
  varname <- as_label(mapping$x)
  ggplot(data, mapping) +
    geom_histogram(bins = BINS, fill = "#9FD0E4", color = "white", linewidth = 0.25) +
    annotate("text", x = -Inf, y = Inf,
             hjust = -0.03, vjust = 1.25,
             label = labels_pretty[varname],
             size = 1.2, fontface = "bold") +
    theme_minimal(base_size = 7) +
    theme(
      panel.grid = element_blank(),
      axis.title = element_blank(),
      axis.text  = element_blank(),
      axis.ticks = element_blank()
    )
}

# --------------------------
# Build and tighten plot
# --------------------------
p <- ggpairs(
  dat,
  columns = match(vars, colnames(dat)),
  upper = list(continuous = wrap(upper_beta)),
  lower = list(continuous = wrap(lower_scatter_with_stars)),
  diag  = list(continuous = wrap(diag_hist_label)),
  labeller = as_labeller(labels_pretty)
)

p <- p +
  theme(
    strip.text    = element_blank(),
    panel.border  = element_rect(color = "grey40", fill = NA, linewidth = 0.3),
    panel.spacing = unit(0.14, "lines"),
    plot.title    = element_blank()
  )

p


```

### 6.3. IOS

```{r, message = FALSE}

# fear
tab_model(lm(inclusionOtherSelf_mean_T2 ~ age + Male + Other + YesCollegeDegree + GroupID + stateEmp_fear, data = dfS_noT3), 
          show.df = T, 
          show.aic = T, 
          show.se = T, 
          show.stat = T, 
          show.std = T)

tab_model(lm(IOS_strangerNum_T2 ~ age + Male + Other + YesCollegeDegree + GroupID + stateEmp_fear, data = dfS_noT3), 
          show.df = T, 
          show.aic = T, 
          show.se = T, 
          show.stat = T, 
          show.std = T)

tab_model(lm(IOS_studyPartner ~ age + Male + Other + YesCollegeDegree + GroupID + stateEmp_fear, data = dfS_noT3), 
          show.df = T, 
          show.aic = T, 
          show.se = T, 
          show.stat = T, 
          show.std = T)


# pain rating 
tab_model(lm(inclusionOtherSelf_mean_T2 ~ age + Male + Other + YesCollegeDegree + GroupID + stateEmp_painRating, data = dfS_noT3), 
          show.df = T, 
          show.aic = T, 
          show.se = T, 
          show.stat = T, 
          show.std = T)

tab_model(lm(IOS_strangerNum_T2 ~ age + Male + Other + YesCollegeDegree + GroupID + stateEmp_painRating, data = dfS_noT3), 
          show.df = T, 
          show.aic = T, 
          show.se = T, 
          show.stat = T, 
          show.std = T)

tab_model(lm(IOS_studyPartner ~ age + Male + Other + YesCollegeDegree + GroupID + stateEmp_painRating, data = dfS_noT3), 
          show.df = T, 
          show.aic = T, 
          show.se = T, 
          show.stat = T, 
          show.std = T)

# unpleasantness of pain 
tab_model(lm(inclusionOtherSelf_mean_T2 ~ age + Male + Other + YesCollegeDegree + GroupID + stateEmp_painUnpleasant, data = dfS_noT3), 
          show.df = T, 
          show.aic = T, 
          show.se = T, 
          show.stat = T, 
          show.std = T)

tab_model(lm(IOS_strangerNum_T2 ~ age + Male + Other + YesCollegeDegree + GroupID + stateEmp_painUnpleasant, data = dfS_noT3), 
          show.df = T, 
          show.aic = T, 
          show.se = T, 
          show.stat = T, 
          show.std = T)


tab_model(lm(IOS_studyPartner ~ age + Male + Other + YesCollegeDegree + GroupID + stateEmp_painUnpleasant, data = dfS_noT3), 
          show.df = T, 
          show.aic = T, 
          show.se = T, 
          show.stat = T, 
          show.std = T)


```

### 6.4. BF 

```{r, message = FALSE}
library(BayesFactor)
library(dplyr)

# Compute BF from t-value
compute_BF_from_t <- function(t, n) {
  bf10 <- BayesFactor::ttest.tstat(t = t, n1 = n, simple = TRUE)
  return(as.numeric(bf10))
}

# Interpret BF
interpret_BF <- function(bf10) {
  if (bf10 < 1/10) return("Strong evidence for H₀")
  if (bf10 < 1/3)  return("Moderate evidence for H₀")
  if (bf10 < 1)    return("Anecdotal evidence for H₀")
  if (bf10 < 3)    return("Anecdotal evidence for H₁")
  if (bf10 < 10)   return("Moderate evidence for H₁")
  return("Strong evidence for H₁")
}

# Outcomes and predictors
outcomes <- c("traitEmpathyIRI_EC_mean_T2", "traitEmpathyIRI_PT_mean_T2",
              "socialConnectedness_mean_T2", "loneliness_mean_T2",
              "inclusionOtherSelf_mean_T2", "IOS_strangerNum_T2", "IOS_studyPartner")
predictors <- c("stateEmp_fear", "stateEmp_painRating", "stateEmp_painUnpleasant")

# Results holder
results_df <- data.frame()

# Loop over models
for (outcome in outcomes) {
  for (predictor in predictors) {
    
    # Fit model
    formula <- as.formula(paste0(outcome, " ~ age + Male + Other + YesCollegeDegree + GroupID + ", predictor))
    model <- lm(formula, data = dfS_noT3)
    model_summary <- summary(model)
    n <- nobs(model)
    
    # Get all coefficients
    coefs <- rownames(model_summary$coefficients)

    # Terms to evaluate
    terms_to_check <- c(predictor)
    
    # Loop over both terms
    for (term in terms_to_check) {
      if (term %in% coefs) {
        t_val <- model_summary$coefficients[term, "t value"]
        bf10 <- compute_BF_from_t(t_val, n)
        bf01 <- 1 / bf10
        evidence <- interpret_BF(bf10)
        
        results_df <- rbind(results_df, data.frame(
          Outcome = outcome,
          Predictor = term,
          t_value = round(t_val, 2),
          N = n,
          BF10 = round(bf10, 3),
          BF01 = round(bf01, 3),
          Evidence = evidence
        ))
      }
    }
  }
}

# Output
print(results_df)

```

## 7. Neuro

### 7.1. Prep data for AFNI analysis

```{r, message = FALSE}
# List of variables to mean center for 3dMVM analysis in AFNI where these are our predictors of pattern similarity 
variables_to_center <- c("loneliness_mean_T2", "traitEmpathyIRI_EC_mean_T2", "traitEmpathyIRI_PT_mean_T2", "stateEmp_fear", "stateEmp_painUnpleasant", "stateEmp_painRating" 
)

# Loop over each variable to mean-center
for (var in variables_to_center) {
  # Check if the column exists in dfS_noT3
  if (var %in% colnames(dfS_noT3)) {
    # Create the mean-centered column
    dfS_noT3[[paste0("mc_", var)]] <- dfS_noT3[[var]] - mean(dfS_noT3[[var]], na.rm = TRUE)
  } else {
    warning(paste("Column", var, "is missing in dfS_noT3 and will be skipped."))
  }
}

# Verify the result
head(dfS_noT3)

# Example of selecting and printing specific columns
mc_df = dfS_noT3 %>%
  select(SubID, GroupID, mc_loneliness_mean_T2, mc_traitEmpathyIRI_EC_mean_T2,mc_traitEmpathyIRI_PT_mean_T2, mc_stateEmp_fear, mc_stateEmp_painUnpleasant, mc_stateEmp_painRating) %>%
  print()


```


### 7.2. AI and dACC Plots based on 3dMVM

```{r, message = FALSE}
library(ggplot2)
library(ggthemes)  # For APA theme
library(ggExtra)
library(papaja)

# List of DVs and their corresponding IVs
dv_iv_mapping <- list(
"EC_patternsim_pain_dACC_MainEffect" = "mc_traitEmpathyIRI_EC_mean_T2",
"mc_stateEmp_fear_patternsim_anticipation_rightAI_Interaction" = "mc_stateEmp_fear",
"mc_stateEmp_painRating_patternsim_pain_leftAI_MainEffect" = "mc_stateEmp_painRating",
"mc_stateEmp_painUnpleasant_patternsim_anticipation_dACC_Interaction" = "mc_stateEmp_painUnpleasant"
)

# Function to generate custom labels for IVs, DVs, and title components
get_labels <- function(iv, dv) {
  iv_labels <- list(
    "mc_traitEmpathyIRI_EC_mean_T2" = "Empathic Concern",
    "mc_loneliness_mean_T2" = "Loneliness",
    "mc_stateEmp_fear" = "State Empathy (Fear)",
    "mc_stateEmp_painUnpleasant" = "State Empathy (Unpleasantness)",
    "mc_traitEmpathyIRI_PT_mean_T2" = "Perspective Taking",
    "mc_stateEmp_painRating" = "State Empathy (Pain Rating)"
  )
  
  dv_labels <- c(
    "decoding_anticipation" = "AUC (Fearful Anticipation)",
    "decoding_pain" = "AUC (Pain)",
    "patternsim_pain" = "Pattern Similarity (Pain)",
    "patternsim_anticipation" = "Pattern Similarity (Fearful Anticipation)",
    "self_pain" = "Brain Activity (Experience Pain)",
    "self_anticipation" = "Brain Activity (Experience Fearful Anticipation)",
    "other_pain" = "Brain Activity (Observe Pain)",
    "other_anticipation" = "Brain Activity (Observe Fearful Anticipation)"
  )
  
  # Extract the DV type (e.g., "decoding_anticipation") and region (e.g., "dACC")
  dv_type <- sub(".*_(decoding_anticipation|decoding_pain|patternsim_pain|patternsim_anticipation|self_pain|self_anticipation|other_pain|other_anticipation).*", "\\1", dv)
  region <- ifelse(grepl("dACC", dv), "dACC",
                   ifelse(grepl("leftAI", dv), "Left AI", "Right AI"))
  
  # Extract effect type (Main Effect, Interaction, or LKM vs PMR)
  effect_type <- ifelse(grepl("MainEffect", dv), "Main Effect",
                        ifelse(grepl("Interaction", dv), "Interaction", "LKM vs PMR"))
  
  iv_label <- iv_labels[[iv]]
  dv_label <- paste(dv_labels[[dv_type]], region)
  
  list(iv_label = iv_label, dv_label = dv_label, effect_type = effect_type)
}

# Loop through each DV-IV pair and create a plot
for (dv in names(dv_iv_mapping)) {
  iv <- dv_iv_mapping[[dv]]
  
  # Check if DV exists in dfS_noT3
  if (dv %in% colnames(dfS_noT3) & iv %in% colnames(dfS_noT3)) {
    labels <- get_labels(iv, dv)
    
    # Create scatter plot with interaction
    plot_interaction <- ggplot(dfS_noT3, aes_string(x = iv, y = dv, color = "GroupID")) +
      geom_point() +
      geom_smooth(method = "lm", aes(fill = "GroupID"), se = TRUE) +
      scale_color_manual(values = c("LKM" = "darkblue", "PMR" = "steelblue1")) +
      scale_fill_manual(values = c("LKM" = "darkblue", "PMR" = "steelblue1"), guide = FALSE) +
      labs(
        x = labels$iv_label,
        y = labels$dv_label,
        color = "Group",
        title = paste(labels$effect_type)
      ) +
      theme_apa()
          # Add marginal density plots for each axis, with color by GroupID
      plot_interaction <- ggMarginal(
        plot_interaction,
        type = "density",
        groupFill = TRUE,  # Color densities by group
        margins = "both",  # Add marginal plots on both axes
        size = 5,  # Adjust marginal plot size
        alpha = 0.4  # Transparency for density plots
      )
          
    # Print the plot
    print(plot_interaction)
    
    # # Save the plot as a file
    # ggsave(
    #   filename = paste0("/my/path/data/plot_", dv, "_vs_", iv, ".png"),
    #   plot = plot_interaction,
    #   width = 8, height = 6
    # )
    
    print(paste("Plot saved for DV:", dv, "and IV:", iv))
  } else {
    print(paste("Skipping DV:", dv, "or IV:", iv, "as it is not found in the dataframe."))
  }
}


```

## 8. Calculate BF for neuro results

### 8.1. Calculate BF for 2sample t-tests

BF01 = 3 means your data are 3x more likely under H0 than H1 — for that specific test. 

```{r, message = FALSE}
library(BayesFactor)
library(dplyr)
library(readr)

# Load CSV
df <- read.csv(file.path(path, "clean/LKMminusControl_mean_Tstats.csv"))

# Degrees of freedom (from AFNI)
df_val <- 52

# Convert t-values to Bayes Factors
results_df <- df %>%
  mutate(
    t_value = MeanTstat_LKMminusControl,
    log_bf10 = sapply(t_value, function(t) ttest.tstat(t = t, n1 = df_val + 2, rscale = 0.707)$bf),
    bf10 = exp(log_bf10),
    bf01 = 1 / bf10,
    evidence_label = case_when(
      bf01 >= 10 | bf01 <= 1/10 ~ "strong evidence for H0",
      bf01 >= 3  | bf01 <= 1/3  ~ "moderate evidence for H0",
      TRUE                      ~ "anecdotal evidence for H0"
    )
  ) %>%
  select(TestName, MaskUsed, t_value, bf01, bf10, evidence_label) %>%
  mutate(across(c(t_value, bf01, bf10), ~ round(.x, 3)))

# View final result
print(results_df)

# Optionally save to CSV
# write.csv(results_df, file.path(path, "LKMminusControl_BayesFactors.csv"), row.names = FALSE)


```

### 8.1. Calculate BF for 3dMVM

```{r, message = FALSE}

library(BayesFactor)
library(dplyr)
library(tidyr)
library(readr)

# Load your data
df <- read.csv(file.path(path, "clean/3dMVM_tstats_summary.csv"))

# Degrees of freedom for all tests (from AFNI output)
df_val <- 50

# Convert t-values to BF for all three effects
df_long <- df %>%
  pivot_longer(cols = starts_with("T_"),
               names_to = "Effect",
               values_to = "t_value") %>%
  mutate(
    # Compute log BF10 from t-values
    log_bf10 = sapply(t_value, function(t) ttest.tstat(t = t, n1 = df_val + 1, rscale = 0.707)$bf),
    bf10 = exp(log_bf10),
    bf01 = 1 / bf10,
    evidence_label = case_when(
      bf01 >= 10 | bf01 <= 1/10 ~ "strong evidence for H0",
      bf01 >= 3  | bf01 <= 1/3  ~ "moderate evidence for H0",
      TRUE                      ~ "anecdotal evidence for H0"
    )
  ) %>%
  select(FileName, MaskUsed, Effect, t_value, bf01, bf10, evidence_label) %>%
  mutate(across(c(t_value, bf01, bf10), ~ round(.x, 3)))

# View result
print(df_long)

# Optionally save
# write.csv(df_long, file.path(path, "MVM_BayesFactors_long.csv"), row.names = FALSE)
```
